---
title: "modelling_llm_performances"
author: "Andrea Nasuto"
date: "2025-04-08"
output: html_document
---

Build a model to assess the performance of a series of LLMs across multiple languages, translations, test/train dataset

The dataset to perform this should have this structure. All variables should be factor/characters:

ID | Model | Language | Translation | Test/Train | Correct
100100 | FT-Multi | IT | 0 | Train | 1
100100 | FT-EN | IT | 1 | Test | 0
100100 | FT-EN-ES | IT | 1 | Train | 1
120300 | LLM2 | EN | EN | Train | 1
...


```{r}
# Set the option to avoid scientific notation
options(scipen = 999)
```

```{r}
# load libraries
library(dplyr)
library(readr)
library(stringr)
library(caret)
library(readxl)
library(pROC)
```

```{r}
# dataset paths
base_path_test <- "/Users/Andrea/Library/CloudStorage/OneDrive-TheUniversityofLiverpool/PhD/code/global-immigration-sentiment/validation/test_dataset_raw"
base_path_train <- "/Users/Andrea/Library/CloudStorage/OneDrive-TheUniversityofLiverpool/PhD/code/global-immigration-sentiment/validation/train_dataset_raw"

output_dir <- "/Users/Andrea/Library/CloudStorage/OneDrive-TheUniversityofLiverpool/PhD/code/global-immigration-sentiment/output/model_performances"

path_translation_test <- "/Users/Andrea/Library/CloudStorage/OneDrive-TheUniversityofLiverpool/PhD/code/global-immigration-sentiment/validation/tweet_translation_checked/test"

path_translation_train <- "/Users/Andrea/Library/CloudStorage/OneDrive-TheUniversityofLiverpool/PhD/code/global-immigration-sentiment/validation/tweet_translation_checked/train"

#path_translation_checked_test <- "/Users/Andrea/Library/CloudStorage/OneDrive-TheUniversityofLiverpool/PhD/code/global-immigration-sentiment/validation/tweet_translation_checked/test/"

#path_translation_checked_train <- "/Users/Andrea/Library/CloudStorage/OneDrive-TheUniversityofLiverpool/PhD/code/global-immigration-sentiment/validation/tweet_translation_checked/train/"
```

```{r}
# Define dataset directories
dirs <- list(
  list(path = base_path_train, type = "Train"),
  list(path = base_path_test, type = "Test"),
  list(path = path_translation_train, type = "Train"),
  list(path = path_translation_test, type = "Test")
)

# Get all file paths for processing
get_all_files <- function(dirs) {
  all_files <- list()
  for (dir in dirs) {
    if (!is.character(dir$path) || length(dir$path) != 1) {
      stop("Each dir in 'dirs' must contain a character path. Found: ", dir$path)
    }

    # List .csv and .xlsx files
    files <- list.files(dir$path, full.names = TRUE, pattern = "\\.(csv|xlsx)$")

    # Exclude files with 'translation' in the name for base_path_train and base_path_test
    if (dir$path %in% c(base_path_train, base_path_test)) {
      files <- files[!grepl("translation", basename(files), ignore.case = TRUE)]
    }
    
    # Exclude files with 'full' in the name
    files <- files[!grepl("full", basename(files), ignore.case = TRUE)]

    for (f in files) {
      all_files[[length(all_files) + 1]] <- list(file = f, type = dir$type)
    }
  }
  return(all_files)
}
```

```{r}
extract_model_dataset <- function(filename) {
  filename <- gsub(".csv", "", filename)
  parts <- unlist(strsplit(filename, "_"))
  
  model_end <- max(grep("llama", parts))
  dataset_start <- min(grep("mig-", parts))
  
  model <- paste(parts[1:model_end], collapse = "-")
  
  dataset_code <- parts[dataset_start]
  dataset <- ifelse(dataset_code %in% names(language_map), language_map[[dataset_code]], dataset_code)
  
  return(data.frame(Model = model, Language = dataset, stringsAsFactors = FALSE))
}
```


```{r}
# language mapping
language_map <- list(
  "mig-ar" = "Arabic",
  "mig-ar-L3" = "Arabic",
  "mig-es" = "Spanish",
  "mig-es-L3" = "Spanish",
  "mig-en" = "English",
  "mig-en-L3" = "English",
  "mig-pt" = "Portuguese",
  "mig-pt-L3" = "Portuguese",
  "mig-fr" = "French",
  "mig-fr-L3" = "French",
  "mig-it" = "Italian",
  "mig-it-L3" = "Italian",
  "mig-de" = "German",
  "mig-de-L3" = "German",
  "mig-hi" = "Hindi",
  "mig-hi-L3" = "Hindi",
  "mig-en-es" = "Eng-Spa",
  "mig-en-es-L3" = "Eng-Spa",
  "mig-multi" = "Multi",
  "mig-multi-L3" = "Multi",
  "mig-hu" = "Hungarian",
  "mig-hu-L3" = "Hungarian",
  "mig-in" = "Indonesian",
  "mig-in-L3" = "Indonesian",
  "mig-es-translation" = "Spanish Translation",
  "mig-es-translation-L3" = "Spanish Translation",
  "mig-tr" = "Turkish",
  "mig-tr-L3" = "Turkish",
  "mig-pl" = "Polish",
  "mig-pl-L3" = "Polish"
)
```

```{r}
# Helper function to read and process files
process_file <- function(file_path, split_type) {
  # Skip files with "full" or "es-translation"
  if (grepl("full", basename(file_path)) || grepl("es-translation", basename(file_path))) {
    return(NULL)
  }
  
  # Skip files that don't contain "llama-32-3B"
  if (!grepl("llama-32-3B", basename(file_path))) {
    return(NULL)
  }
  
  # Read the file based on its extension (CSV or Excel)
  file_extension <- tools::file_ext(file_path)
  if (file_extension == "csv") {
    df <- read_csv(file_path, col_types = cols(.default = "c"))
  } else if (file_extension %in% c("xlsx", "xls")) {
    df <- read_excel(file_path)
  } else {
    return(NULL)  # Skip unsupported file types
  }
  
  # If the data frame is empty or null, return NULL
  if (is.null(df) || nrow(df) == 0) return(NULL)

  # Extract model and language information from filename
  meta <- extract_model_dataset(basename(file_path))

# Set Translation value based on the `translation_check` column in the dataset
  if ("translation_check" %in% colnames(df)) {
    df$TranslationQuality <- ifelse(df$translation_check == 1, 'Good', 'Bad')
    df$Translation <- 1
    #print(paste("Processed translation_checked file:", file_path))
  } else if (startsWith(file_path, path_translation_train) || startsWith(file_path, path_translation_test)) {
    # If file is from path_translation_train or path_translation_test, set Translation to 1
    df$Translation <- 1
    df$TranslationQuality <- 'Unknown'
  } else {
    # Otherwise, set Translation to 0
    df$Translation <- 0
    df$TranslationQuality <- 'Not Translated'
  }
  
  # Process the dataset
  df_processed <- df %>%
    mutate(
      Model = meta$Model,
      Language = meta$Language,
      Train_Test = split_type,
      Correct = ifelse(label_llama == label, 1, 0),
      file_path = file_path
    ) %>%
    select(message_id, Model, Language, Translation, TranslationQuality, Train_Test, Correct, label, label_llama, plaintext, file_path)

  return(df_processed)
}
```

```{r}
assign_new_ids <- function(df) {
  df %>%
    group_by(plaintext) %>%
    mutate(newId = cur_group_id()) %>%
    ungroup()
}
```

### Create dataset

```{r}
# Combine the processed data from all files
all_files <- get_all_files(dirs)
list_of_processed_files <- lapply(all_files, function(file_info) {
  process_file(file_info$file, file_info$type)
})
```

```{r}
# create dataset
df <- do.call(rbind, Filter(Negate(is.null), list_of_processed_files))
```

```{r}
# assign new ID
df <- assign_new_ids(df)
```

```{r}
# include only quantized at 4bit models
df <- df %>%
  filter(grepl("llama-32-3B", Model) & grepl("Q4", Model))
```

```{r}
# remove duplicate - each observation should have a unique pair of newId and filepath
df <- df %>%
  distinct(file_path, newId, .keep_all = TRUE) 
```

```{r}
# rename models after filtering
model_map <- function(model) {
  if (grepl("llama-32-3B-en-Q4-translation", model)) {
    return("English-Translation")
  } else if (grepl("llama-32-3B-en-es", model)) {
    return("English-Spanish")
  } else if (grepl("llama-32-3B-multi", model)) {
    return("Multilanguage")
  } else if (grepl("llama-32-3B-es", model)) {
    return("Spanish")
  } else if (grepl("llama-32-3B-en", model)) {
    return("English")
  } else {
    return(model)
  }
}
```


```{r}
# Rename models after filtering in df
df$Model <- sapply(df$Model, model_map)
```

```{r}
# remove values where Language is Multi
df <- df %>%
  filter(!(Language %in% c("Multi", "Eng-Spa")))
```

```{r}
# remove all observation where label is NA
df <- df %>%
  filter(!is.na(label))
```

```{r}
# Ensure all variables are factors (if needed)
df <- df %>%
  mutate(
    message_id = as.character(message_id),
    Model = as.factor(Model),
    Language = as.factor(Language),
    Train_Test = as.factor(Train_Test),
    Translation = as.factor(Translation),
    TranslationQuality = as.factor(TranslationQuality),
    Correct = as.factor(Correct)
  )
```

### Checks
```{r}
# check if there are NAs in df
na_count <- sapply(df, function(x) sum(is.na(x)))
print(na_count)

# check how many correct values are in Correct
table(df$Correct)
```

```{r}
# count number of obs per language
df %>%
  count(Language, name = "n_observations") %>%
  arrange(desc(n_observations))

table(df$Language)
```
```{r}
# count the number of obs with Translation == 2
translation_count <- df %>%
  filter(TranslationQuality == 'Good') %>%
  group_by(Model, Language) %>%
  summarise(count = n())
translation_count
```

## Model
```{r}
# Set reference levels for the factors
df$Model <- relevel(df$Model, ref = "English") #llama-32-3B-en-Q4 - all the models are 4bit
df$Language <- relevel(df$Language, ref = "English")
df$Train_Test <- relevel(df$Train_Test, ref = "Train")
df$label <- factor(df$label)
df$label <- relevel(df$label, ref = "neutral")
df$TranslationQuality <- factor(df$TranslationQuality)
df$TranslationQuality <- relevel(df$TranslationQuality, ref = "Not Translated") # Not Translated
df$Translation <- factor(df$Translation)

# Fit the logistic regression model
logit_model <- glm(Correct ~ Model + Language + label + Train_Test + TranslationQuality, # add Translated not Translated and Translated(0,1)*TranslationQuality(unknown, bad, good)  
                   data = df, 
                   family = binomial())

# Summarize the model
summary(logit_model)
```

```{r}
# Fit the logistic regression model
logit_model_2 <- glm(Correct ~ Model*Language + label + Train_Test + TranslationQuality, # add Translated not Translated and Translated(0,1)*TranslationQuality(unknown, bad, good)  
                   data = df, 
                   family = binomial())

# Summarize the model
summary(logit_model_2)
```

##### Random Forest
```{r}
# Load the package
library(randomForest)

# Fit a random forest model
#rf_model <- randomForest(Correct ~ Model + Language + Train_Test + TranslationQuality + label, 
#                         data = df, 
#                         importance = TRUE, 
#                         ntree = 500)

# Summarize the random forest model
#print(rf_model)

# Plot the importance of the variables
#importance(rf_model)  # This will show the importance of each variable
#varImpPlot(rf_model)  # This will generate a plot showing the importance of each variable
```
```{r}
library(rpart)
library(rpart.plot)
tree <- rpart(Correct ~ Model + Language + Train_Test + TranslationQuality + label, # add TrainSize
                         data = df)
rpart.plot(tree, yesno = 2)
```

