---
title: "modelling_llm_performances"
author: "Andrea Nasuto"
date: "2025-04-08"
output: html_document
---

This document assembles a message-by-model evaluation dataset and uses it to compare classification performance across model families, source languages, translation status, and train/test splits. The workflow has two stages:

1. It reads raw and translated classification outputs, standardises their metadata, and combines them into a single modelling table.
2. It estimates a series of logistic models and descriptive summaries to show where accuracy changes across models, languages, and label compositions.

The modelling dataset is organised at the `message_id × model output` level, so a single underlying tweet can appear more than once when it has been classified by multiple models or under both translated and non-translated conditions. After ingestion, the notebook derives a text-based shared identifier (`newId`) to link repeated observations of the same plaintext message across files.

The main variables used in the analysis are:

- `message_id`: original identifier provided in the source file
- `newId`: harmonised identifier created from the plaintext tweet so the same message can be tracked across models
- `plaintext`: original tweet text used for matching and modelling
- `Model`: model family that produced the classification, standardised to labels such as `English`, `Spanish`, `English-Spanish`, and `Multilanguage`
- `Language`: language of the tweet content
- `Train_Test`: whether the observation comes from the training or test evaluation split
- `Translation`: binary flag indicating whether the model classified a translated version of the tweet
- `TranslationQuality`: status of the translated text, coded as `Not Translated`, `Unknown`, `Good`, or `Bad`
- `label`: human-coded gold label
- `label_llama`: model-predicted label
- `Correct`: binary outcome equal to `1` when `label_llama == label` and `0` otherwise

An illustrative schema is shown below:

message_id | newId | Model | Language | Train_Test | Translation | TranslationQuality | label | label_llama | Correct
100100 | 1 | English | Italian | Train | 0 | Not Translated | neutral | neutral | 1
100100 | 1 | English-Spanish | Italian | Train | 1 | Good | neutral | anti-immigration | 0
100100 | 1 | Multilanguage | Italian | Test | 0 | Not Translated | neutral | neutral | 1
120300 | 2 | Spanish | English | Test | 1 | Unknown | unrelated | unrelated | 1
...


```{r}
# set options
options(scipen = 999)
```

```{r}
# load libraries
library(dplyr)
library(ggplot2)
library(purrr)
library(readr)
library(stringr)
library(broom)
library(caret)
library(readxl)
library(rpart)
library(rpart.plot)
library(viridis)
library(showtext)
library(sysfonts)
library(forcats)
library(scales)
library(lme4)

# load fonts and helper functions
base_family <- "sans"

try({
  font_add_google("Roboto Condensed", "roboto_condensed")
  showtext_auto()
  base_family <- "roboto_condensed"
}, silent = TRUE)

source("/Users/Andrea/Library/CloudStorage/OneDrive-TheUniversityofLiverpool/PhD/code/immigration-llm/model_performance/helpers.R")
```

```{r}
# define project paths once so the rest of the notebook can reuse them
project_root <- "/Users/Andrea/Library/CloudStorage/OneDrive-TheUniversityofLiverpool/PhD/code/immigration-llm"
knitr::opts_knit$set(root.dir = project_root)

base_path_test <- file.path(project_root, "model_performance", "data", "test_dataset_raw")
base_path_train <- file.path(project_root, "model_performance", "data", "train_dataset_raw")
path_translation_test <- file.path(project_root, "model_performance", "data", "test_dataset_translation")
path_translation_train <- file.path(project_root, "model_performance", "data", "train_dataset_translation")
output_dir <- file.path(project_root, "model_performance", "output")
```

## Data Preparation

```{r}
# ingest raw and translated evaluation outputs into one long table
df <- build_combined_dataset(
  raw_test_dir = base_path_test,
  raw_train_dir = base_path_train,
  translation_test_dir = path_translation_test,
  translation_train_dir = path_translation_train
) %>%
  prepare_modelling_dataset() %>%
  add_share_unrelated() %>%
  add_language_share()
```

## Descriptive Statistics

```{r}
# inspect missingness and outcome balance
na_count <- sapply(df, function(x) sum(is.na(x)))
print(na_count)
table(df$Correct)
```

```{r}
# summarise observations by language and model
df %>%
  count(Language, name = "n_observations") %>%
  arrange(desc(n_observations))

table(df$Language, df$Model)
```

```{r}
# summarise unique tweets by language
df_unique_counts <- df %>%
  distinct(newId, Language) %>%
  count(Language, name = "n_unique_observations") %>%
  arrange(desc(n_unique_observations))

df_unique_counts
```

```{r}
# summarise checked translations
translation_count <- df %>%
  filter(TranslationQuality == "Good") %>%
  group_by(Model, Language) %>%
  summarise(count = n(), .groups = "drop")

translation_count
```

```{r}
# create and save the label distribution table
label_distribution <- df %>%
  group_by(newId) %>%
  summarise(label = first(label), .groups = "drop") %>%
  count(label, name = "n_tweets") %>%
  mutate(share_total_tweets = n_tweets / sum(n_tweets)) %>%
  arrange(desc(n_tweets))

write_csv(label_distribution, file.path(output_dir, "label_distribution.csv"))
label_distribution
```

## Models

### Model 1
```{r}
# use consistent baselines so coefficients are interpretable across models
df <- set_reference_levels(df)

# fit the logistic regression model
baseline_glm <- glm(
  Correct ~ Model + Language + Train_Test + TranslationQuality + label,
  data = df,
  family = binomial()
)

# summarise the model
summary(baseline_glm)
```

```{r}
# prepare coefficient output for the baseline coefficient plot
baseline_results <- tidy_baseline_terms(baseline_glm)
```

```{r}
# set the colour palette
cb_palette <- c(
  "#E69F00", "#56B4E9", "#009E73", "#F0E442",
  "#0072B2", "#D55E00", "#CC79A7", "#999999"
)
```

```{r}
# define the named palette
cb_palette_named <- c(
  "Model"       = cb_palette[1],
  "Language"    = cb_palette[5],
  "Label"       = cb_palette[3],
  "Translation" = cb_palette[6]
)

baseline_plot <- baseline_results %>%
  mutate(
    clean_term = fct_reorder(clean_term, estimate),
    variable_group = fct_reorder(variable_group, abs_estimate, .fun = median)
  ) %>%
  ggplot(aes(x = estimate, y = clean_term, color = variable_group)) +
  geom_vline(xintercept = 0, linetype = "dashed", alpha = 0.7, linewidth = 0.5) +
  geom_errorbarh(aes(xmin = conf.low, xmax = conf.high),
                 height = 0.4, alpha = 1, linewidth = 0.6) +
  geom_point(size = 2.0, alpha = 1) +
  facet_grid(rows = vars(variable_group), scales = "free_y", space = "free_y", switch = "y") +
  scale_color_manual(values = cb_palette_named, guide = "none") +
  labs(
    x = "Coefficient Estimates", y = NULL
  ) +
  theme(
    text              = element_text(family = base_family),
    strip.text.y.left = element_text(size = 11, face = "bold"),
    axis.text.y       = element_text(size = 12, lineheight = 0.8),
    axis.text.x       = element_text(size = 10),
    axis.title.x    = element_text(size = 14, margin = margin(t = 12)),
    panel.grid.minor  = element_blank(),
    strip.placement   = "outside",
    plot.margin       = margin(10, 30, 10, 10)
  )

baseline_plot

# save the plot
ggsave(filename = "model1_baseline_plot.png", plot = baseline_plot, path = output_dir, width = 10, height = 6, dpi = 300)
```


### Model 2
```{r}
# estimate whether model performance changes by language
model_language_glm <- glm(
  Correct ~ Model * Language + label + Train_Test + TranslationQuality,
  data = df,
  family = binomial()
)

# summarise the model
summary(model_language_glm)
```

```{r}
# isolate the model and language terms for the interaction plot
model2_results <- tidy_model_language_terms(model_language_glm)

# plot model 2 coefficients
model_language_coeff_plot <- model2_results %>%
  mutate(
    clean_term = fct_reorder(clean_term, estimate),
    variable_group = fct_reorder(variable_group, abs_estimate, .fun = median)
  ) %>%
  ggplot(aes(x = estimate, y = clean_term, color = variable_group)) +
  geom_vline(xintercept = 0, linetype = "dashed", alpha = 0.7, linewidth = 0.5) +
  geom_errorbarh(aes(xmin = conf.low, xmax = conf.high), height = 0.2, alpha = 0.7, linewidth = 0.6) +
  geom_point(size = 1.5, alpha = 1) +
  facet_wrap(~variable_group, scales = "free_y", ncol = 1, strip.position = "left") +
  scale_color_manual(values = cb_palette, guide = "none") +
  labs(
    x = "Log-odds (Coefficient Estimate)", y = NULL
  ) +
  theme(
    text = element_text(family = base_family),
    plot.title = element_text(size = 16, face = "bold"),
    plot.subtitle = element_text(size = 12, color = "grey40"),
    strip.text = element_text(size = 9, face = "bold"),
    axis.text.y = element_text(size = 4),
    axis.text.x = element_text(size = 10),
    panel.grid.minor = element_blank()
  )

# show the plot
model_language_coeff_plot
```


```{r}
# convert model coefficients into a full model-language prediction surface
combinations <- compute_model_language_surface(model_language_glm, df)
```

```{r}
combinations$Model <- factor(
  combinations$Model,
  levels = c("English", "Spanish", "English-Spanish", "Multilanguage")
)

# order languages by average predicted accuracy
accuracy_grouped_combination <- combinations %>%
  group_by(Language) %>%
  summarise(mean_accuracy = mean(accuracy, na.rm = TRUE)) %>%
  arrange(mean_accuracy)

combinations$Language <- factor(
  combinations$Language,
  levels = accuracy_grouped_combination$Language
)
```

```{r}
# plot heatmaps for model 2
heatmap_logodds <- ggplot(combinations, aes(x = Model, y = Language, fill = log_odds)) +
  geom_tile(color = "white") +
  scale_fill_viridis_c(
    option = "viridis",
    direction = 1,
    name = "Log-odds"
  ) +
  theme_minimal() +
  theme(
    text = element_text(family = base_family),
    plot.title = element_text(family = base_family, face = "bold"),
    axis.text = element_text(family = base_family),
    legend.title = element_text(family = base_family),
    legend.text = element_text(family = base_family),
    axis.text.y = element_text(size = 11),
    axis.text.x = element_text(size = 11),
    axis.title.x    = element_text(size = 13, margin = margin(t = 12)),
    axis.title.y    = element_text(size = 13)
    )

heatmap_accuracy <- ggplot(combinations, aes(x = Model, y = Language, fill = accuracy)) +
  geom_tile(color = "white") +
  scale_fill_gradient(low = "white", high = "darkgreen", name = "Predicted Accuracy") +
  theme_minimal() +
  theme(
    text = element_text(family = base_family),
    plot.title = element_text(family = base_family, face = "bold"),
    axis.text = element_text(family = base_family),
    legend.title = element_text(family = base_family),
    legend.text = element_text(family = base_family)
  )

# save the heatmaps
ggsave(filename = "model_language_interaction_heatmap_logodds.png", plot = heatmap_logodds, path = output_dir, width = 8, height = 6, dpi = 300)
ggsave(filename = "model_language_interaction_heatmap_accuracy.png", plot = heatmap_accuracy, path = output_dir, width = 8, height = 6, dpi = 300)

# print the plots
print(heatmap_logodds)
print(heatmap_accuracy)
```

### Model 3
```{r}
# estimate whether model performance changes as the share of unrelated tweets rises
share_unrelated_glm <- glm(
  Correct ~ Model * Language + Model * Share_Unrelated + Train_Test + TranslationQuality,
  data = df,
  family = binomial()
)

# summarise the model
summary(share_unrelated_glm)
```

```{r}
# create a prediction grid to visualise how accuracy changes with label composition
df_plot_model3 <- expand.grid(
  Share_Unrelated = seq(min(df$Share_Unrelated, na.rm = TRUE), max(df$Share_Unrelated, na.rm = TRUE), length.out = 100),
  Model = levels(df$Model),
  Language = levels(df$Language),
  Train_Test = "Train",
  TranslationQuality = "Not Translated"
)

df_plot_model3$Accuracy <- predict(share_unrelated_glm, newdata = df_plot_model3, type = "response")
```

```{r}
# set colours for the model 3 line chart
model_cols <- c(
  "English" = "#9A9A9A",
  "Spanish" = "#E69F00",
  "English-Spanish" = "#FE6100",
  "Multilanguage" = "#0072B2"
)

share_unrelated_prediction_plot <- ggplot(df_plot_model3, aes(x = Share_Unrelated, y = Accuracy, color = Model)) +
  geom_line(size = 1.2) +
  facet_wrap(~ Language, ncol = 4) +
  scale_color_manual(values = model_cols) +
  labs(
    x     = "Share of Unrelated Tweets",
    y     = "Mean Accuracy",
    color = "Model"
  ) +
  scale_x_continuous(labels = scales::percent_format(accuracy = 1)) +
  theme_minimal(base_size = 16) +
  theme(
    text             = element_text(family = base_family),
    axis.title       = element_text(size = 18),
    axis.text        = element_text(size = 14),
    strip.text       = element_text(size = 16),
    legend.title     = element_text(size = 16),
    legend.text      = element_text(size = 14),
    legend.position  = "top",
    panel.grid.minor = element_blank()
  )

# save the plot
ggsave(filename = "model_shareunrelated_language_interaction.png", plot = share_unrelated_prediction_plot, path = output_dir, width = 12, height = 8, dpi = 300)
share_unrelated_prediction_plot
```
The interaction specification suggests that the effect of `Share_Unrelated` is not uniform across all languages. In some language panels, the `Multilanguage` model becomes comparatively stronger only when the share of unrelated tweets is very high.

By contrast, the direct `Model × Share_Unrelated` interaction is more stable across languages and indicates that the `Multilanguage` model tends to underperform the `English-Spanish` and `Spanish` models once unrelated tweets make up a large share of the sample.

```{r}
# isolate the coefficients tied to model type and unrelated-share interactions
model3_results <- tidy_share_unrelated_terms(share_unrelated_glm)

# plot model 3 coefficients
share_unrelated_coeff_plot <- model3_results %>%
  mutate(
    clean_term = fct_reorder(clean_term, estimate),
    variable_group = fct_reorder(variable_group, abs_estimate, .fun = median)
  ) %>%
  ggplot(aes(x = estimate, y = clean_term, color = variable_group)) +
  geom_vline(xintercept = 0, linetype = "dashed", alpha = 0.7, linewidth = 0.5) +
  geom_errorbarh(aes(xmin = conf.low, xmax = conf.high), height = 0.2, alpha = 0.7, linewidth = 0.6) +
  geom_point(size = 2.5, alpha = 0.9) +
  facet_wrap(~variable_group, scales = "free_y", ncol = 1, strip.position = "left") +
  scale_color_manual(values = cb_palette, guide = "none") +
  labs(
    x = "Estimates", y = NULL
  ) +
  theme(
    text = element_text(family = base_family),
    plot.title = element_text(size = 16, face = "bold"),
    plot.subtitle = element_text(size = 12, color = "grey40"),
    strip.text = element_text(size = 9, face = "bold"),
    axis.text.y = element_text(size = 9),
    axis.text.x = element_text(size = 10),
    panel.grid.minor = element_blank()
  )

share_unrelated_coeff_plot

# save the plot
ggsave(filename = "model_interaction_language_share_unrelated_plot.png", plot = share_unrelated_coeff_plot, path = output_dir, width = 10, height = 6, dpi = 300)
```


### Model 4


This specification tests whether the `Multilanguage` model still underperforms once the evaluation language is partially standardised through translation. Restricting the sample to translated tweets reduces variation coming from the original source language and makes the comparison closer to a common-language setting.

In this translated subset, the coefficients still suggest weaker performance for the `Multilanguage` model than for the `English-Spanish` and `Spanish` models on `unrelated` content. That pattern is consistent with the idea that the performance gap is not driven only by the original tweet language.

```{r}
# fit the translation-only interaction model
df_translation <- df %>%
  filter(Translation == 1) %>%
  droplevels()

model4_formula <- if (nlevels(df_translation$Model) > 1) {
  Correct ~ Model * label + Language + Train_Test + TranslationQuality
} else {
  Correct ~ label + Language + Train_Test + TranslationQuality
}

translated_label_glm <- glm(
  model4_formula,
  data = df_translation,
  family = binomial()
)

summary(translated_label_glm)
```

```{r}
# keep only the translated-sample terms needed for interpretation
model4_results <- tidy_translation_label_terms(translated_label_glm)
```

```{r}
model4_plot1 <- model4_results %>%
  mutate(clean_term = fct_reorder(clean_term, estimate)) %>%
  ggplot(aes(x = estimate, y = clean_term, color = variable_group)) +
  geom_vline(xintercept = 0, linetype = "dashed", alpha = 0.7, linewidth = 0.5) +
  geom_errorbarh(aes(xmin = conf.low, xmax = conf.high), height = 0.3, alpha = 0.8, linewidth = 0.8) +
  geom_point(size = 3, alpha = 0.9) +
  scale_color_manual(values = cb_palette, name = "Variable Group") +
  labs(
    subtitle = "Log-odds estimates with 95% confidence intervals",
    x = "Coefficient Estimates", y = NULL,
    caption = "Dashed line at zero indicates no effect"
  ) +
  theme(
    plot.subtitle = element_text(size = 12, color = "grey40"),
    axis.text.y = element_text(size = 9),
    axis.text.x = element_text(size = 10),
    axis.title.x = element_text(size = 11),
    legend.position = "bottom",
    legend.title = element_text(size = 10),
    panel.grid.minor = element_blank()
  )
```

```{r}
translated_label_coeff_plot <- model4_results %>%
  mutate(
    clean_term = fct_reorder(clean_term, estimate),
    variable_group = fct_reorder(variable_group, abs_estimate, .fun = median)
  ) %>%
  ggplot(aes(x = estimate, y = clean_term, color = variable_group)) +
  geom_vline(xintercept = 0, linetype = "dashed", alpha = 0.7, linewidth = 0.5) +
  geom_errorbarh(aes(xmin = conf.low, xmax = conf.high), height = 0.2, alpha = 0.7, linewidth = 0.6) +
  geom_point(size = 2.5, alpha = 0.9) +
  facet_wrap(~variable_group, scales = "free_y", ncol = 1, strip.position = "left") +
  scale_color_manual(values = cb_palette, guide = "none") +
  labs(
    x = "Coefficient Estimates", y = NULL
  ) +
  theme_minimal(base_family = base_family) +
  theme(
    strip.text = element_text(size = 11, face = "bold"),
    axis.text.y = element_text(size = 13),
    axis.text.x = element_text(size = 10),
    axis.title.x    = element_text(size = 14, margin = margin(t = 12)),
    panel.grid.minor = element_blank()
  )

# save the plot
ggsave(filename = "model_label_interaction.png", plot = translated_label_coeff_plot, path = output_dir, width = 10, height = 6, dpi = 300)
translated_label_coeff_plot
```

```{r}
model4_plot3 <- model4_results %>%
  mutate(
    clean_term = fct_reorder(clean_term, estimate),
    sig_shape = case_when(
      p.value < 0.001 ~ 19,
      p.value < 0.01 ~ 17,
      p.value < 0.05 ~ 15,
      TRUE ~ 1
    )
  ) %>%
  ggplot(aes(x = estimate, y = clean_term)) +
  geom_vline(xintercept = 0, linetype = "dashed", alpha = 0.7, linewidth = 0.5) +
  geom_segment(aes(x = conf.low, xend = conf.high, y = clean_term, yend = clean_term,
                   color = variable_group), alpha = 0.6, linewidth = 1) +
  geom_point(aes(color = variable_group, shape = factor(sig_shape)), size = 3) +
  scale_color_manual(values = cb_palette, name = "Variable Group") +
  scale_shape_manual(values = c(1, 15, 17, 19),
                     name = "Significance",
                     labels = c("n.s.", "p < 0.05", "p < 0.01", "p < 0.001")) +
  labs(
    x = "Log-odds (Coefficient Estimate)", y = NULL
  ) +
  theme(
    axis.text.y = element_text(size = 9),
    axis.text.x = element_text(size = 10),
    legend.position = "bottom",
    panel.grid.minor = element_blank()
  )
```

```{r}
model4_plot4 <- model4_results %>%
  filter(!variable_group %in% c("Translation Quality", "Language")) %>%
  mutate(readable_label = clean_term) %>%
  mutate(readable_label = fct_reorder(readable_label, estimate)) %>%
  ggplot(aes(y = readable_label, color = variable_group, fill = variable_group)) +
  geom_vline(xintercept = 0, linetype = "solid", alpha = 0.3, linewidth = 1) +
  geom_col(aes(x = estimate), alpha = 0.6, width = 0.6) +
  geom_errorbarh(aes(xmin = conf.low, xmax = conf.high), 
                 height = 0.2, alpha = 0.8, linewidth = 0.8, color = "black") +
  geom_point(aes(x = estimate), size = 1.5, color = "black") +
  scale_fill_manual(values = cb_palette, name = "Variable Type") +
  scale_color_manual(values = cb_palette, name = "Variable Type") +
  labs(
    x = "Coefficient Estimates", y = NULL
  ) +
  theme_minimal(base_family = base_family) +
  theme(
    plot.caption = element_text(size = 9, color = "grey50", hjust = 0),
    axis.text.y = element_text(size = 9, color = "grey20"),
    axis.text.x = element_text(size = 10),
    legend.position = "bottom",
    panel.grid.minor = element_blank()
  )
```

Create heatmaps of Model 4 results across labels and languages

```{r}
# build a language-by-model prediction table within translated observations
languages <- c("Turkish", "Spanish", "Portuguese", "Polish", "Korean", "Italian",
               "Indonesian", "Hungarian", "Hindi", "German", "French", "Arabic")
df_plot_heatmap <- build_translation_heatmap_data(
  model = translated_label_glm,
  languages = languages,
  translation_df = df_translation,
  translation_quality = "Unknown"
)

# plot the heatmap
ggplot(df_plot_heatmap, aes(x = Model, y = Language, fill = Accuracy)) +
  geom_tile(color = "white") +
  geom_text(
    aes(label = sprintf("%.1f", Accuracy * 100)),
    size = 3.2,
    family = base_family,
    color = "white"
  ) +
  facet_wrap(~Label, ncol = 2) +
  scale_fill_viridis(
    name = "Accuracy",
    option = "D",
    labels = scales::percent_format(accuracy = 1)
  ) +
  labs(
    title = "Model 4 Accuracy Across Labels and Languages",
    x = NULL,
    y = NULL
  ) +
  theme(
    legend.position = "right",
    strip.text = element_text(face = "bold"),
    axis.text.x = element_text(angle = 45, hjust = 1),
    axis.text.y = element_text(size = 10),
    axis.title = element_blank(),
    plot.title = element_text(face = "bold", size = 14, margin = margin(b = 10))
  )
```

### Model 5

Understand how pretraining impacts accuracy performances
```{r}
# create the logged language share
df$LogLangShare <- log1p(df$LangShare)

# fit the logistic regression model
logit_model_interact <- glm(Correct ~ Model * LangShare + Language + label + Train_Test + TranslationQuality + label,
                            data = df,
                            family = binomial())

summary(logit_model_interact)
```

```{r}
# convert the outcome to numeric
df$Correct_Num <- as.numeric(as.character(df$Correct))

lang_summary <- df %>%
  group_by(Language, LangShare) %>%
  summarise(accuracy = mean(Correct_Num, na.rm = TRUE), n = n()) %>%
  ungroup()
```

```{r}
# summarise accuracy by language share
df_summary <- df %>%
  filter(Language != "English") %>%
  filter(label != "unrelated") %>%
  group_by(Language, Model, LangShare) %>%
  summarise(accuracy = mean(Correct_Num), .groups = "drop")

# plot accuracy by language and model
ggplot(df_summary, aes(x = LangShare, y = accuracy, color = Model)) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(title = "Accuracy by Language Share and Model Type",
       x = "Language Share in Pretraining (%)",
       y = "Mean Accuracy") +
  theme_minimal()
```

```{r}
# order languages by mean language share
language_order <- df_summary %>%
  group_by(Language) %>%
  summarise(mean_share = mean(LangShare), .groups = "drop") %>%
  arrange(mean_share) %>%
  pull(Language)

df_summary$Language <- factor(df_summary$Language, levels = language_order)
```


```{r}
# plot model 5 summaries
ggplot(df_summary, aes(x = LangShare, y = accuracy, color = Language)) +
  geom_point(size = 2) +
  geom_smooth(method = "lm", se = FALSE, color = "black", aes(group = 1)) +
  facet_wrap(~ Model) +
  labs(
    x = "Share of Language in LLAMA Pretraining",
    y = "Mean Accuracy",
    color = "Language"
  ) +
  theme_minimal(base_size = 16) +
  theme(
    text = element_text(family = base_family),
    axis.title = element_text(size = 18),
    axis.text = element_text(size = 14),
    legend.title = element_text(size = 16),
    legend.text = element_text(size = 14),
    strip.text = element_text(size = 14),
    axis.title.x = element_text(margin = margin(t = 15))
  )

# save the plot
ggsave(file.path(output_dir, "accuracy_vs_language_share_by_model.png"), width = 10, height = 6)
```

```{r}
# summarise model 5 slopes
coefficients_table <- df_summary %>%
  group_by(Model) %>%
  do(tidy(lm(accuracy ~ LangShare, data = .))) %>%
  filter(term == "LangShare") %>%
  select(Model, Estimate = estimate, Std_Error = std.error, p_value = p.value)

print(coefficients_table)
```

## Confusion Matrix and Confidence Intervals
```{r}
# split the non-translation data
df_non_translation <- df %>% filter(Translation == 0)

df_test <- df_non_translation %>% filter(Train_Test == "Test")
df_train <- df_non_translation %>% filter(Train_Test == "Train")
```


```{r}
# generate accuracy statistics
accuracy_test <- generate_accuracy_df(df_test, "Test")
accuracy_train <- generate_accuracy_df(df_train, "Train")
```

```{r}
# save confusion matrix outputs
generate_and_save_cm(df_test, "Test", output_dir)
generate_and_save_cm(df_train, "Train", output_dir)
```

```{r}
# plot test intervals
chart_title <- "Accuracy Confidence Intervals by Language and Dataset (Test)"
plot_accuracy_intervals(accuracy_test, chart_title, output_dir)
```

```{r}
# plot train intervals
chart_title <- "Accuracy Confidence Intervals by Language and Dataset (Train)"
plot_accuracy_intervals(accuracy_train, chart_title, output_dir)
```

## Multilevel Models
```{r}
# fit the multilevel logistic regression
logit_mixed <- glmer(Correct ~ Model * Language + label + Train_Test + TranslationQuality + 
             (1 | Language), 
      data = df,
      family = binomial(),
      nAGQ = 0,
      control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 10000)))

# summarise the model
summary(logit_mixed)
```
```{r}
# fit the random slope multilevel model
multilevel_interaction <- glmer(Correct ~ Model + label + Train_Test + TranslationQuality + 
             (1 + Model | Language), 
      data = df,
      family = binomial(),
      nAGQ = 0,
      control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 10000)))

summary(multilevel_interaction)

```

## Random Forest
```{r}
# fit the random forest style tree model
tree <- rpart(Correct ~ Model + Language + Train_Test + TranslationQuality + label,
              data = df)
rpart.plot(tree, yesno = 2)
```
