---
title: "modelling_llm_performances"
author: "Andrea Nasuto"
date: "2025-04-08"
output: html_document
---

Build a model to assess the performance of a series of LLMs across multiple languages, translations, test/train dataset

The dataset to perform this should have this structure. All variables should be factor/characters:

ID | Model | Language | Translation | Test/Train | Correct
100100 | FT-Multi | IT | 0 | Train | 1
100100 | FT-EN | IT | 1 | Test | 0
100100 | FT-EN-ES | IT | 1 | Train | 1
120300 | LLM2 | EN | EN | Train | 1
...


```{r}
# Set the option to avoid scientific notation
options(scipen = 999)
```

```{r}
# load libraries
library(dplyr)
library(ggplot2)
library(tidyr)
library(purrr)
library(readr)
library(stringr)
library(broom)
library(caret)
library(readxl)
library(pROC)
library(rpart)
library(rpart.plot)
```

```{r}
# set working directory
knitr::opts_knit$set(root.dir = "/Users/Andrea/Library/CloudStorage/OneDrive-TheUniversityofLiverpool/PhD/code/immigration-llm")

# dataset paths
base_path_test <- "/Users/Andrea/Library/CloudStorage/OneDrive-TheUniversityofLiverpool/PhD/code/immigration-llm/model_performance/data/test_dataset_raw"

base_path_train <- "/Users/Andrea/Library/CloudStorage/OneDrive-TheUniversityofLiverpool/PhD/code/immigration-llm/model_performance/data/train_dataset_raw"

path_translation_test <- "/Users/Andrea/Library/CloudStorage/OneDrive-TheUniversityofLiverpool/PhD/code/immigration-llm/model_performance/data/test_dataset_translation"

path_translation_train <- "/Users/Andrea/Library/CloudStorage/OneDrive-TheUniversityofLiverpool/PhD/code/immigration-llm/model_performance/data/train_dataset_translation"

#path_translation_checked_test <- "/Users/Andrea/Library/CloudStorage/OneDrive-TheUniversityofLiverpool/PhD/code/global-immigration-sentiment/validation/tweet_translation_checked/test/"

#path_translation_checked_train <- "/Users/Andrea/Library/CloudStorage/OneDrive-TheUniversityofLiverpool/PhD/code/global-immigration-sentiment/validation/tweet_translation_checked/train/"

# output path
output_dir <- "output"
```

```{r}
# Get all CSV files
test_files <- list.files(base_path_test, pattern = "\\.csv$", full.names = TRUE)
train_files <- list.files(base_path_train, pattern = "\\.csv$", full.names = TRUE)

# Get all CSV files
test_translation_files <- list.files(path_translation_test, pattern = "\\.(csv|xlsx)$", full.names = TRUE)
train_translation_files <- list.files(path_translation_train, pattern = "\\.(csv|xlsx)$", full.names = TRUE)

# Define dataset directories
dirs <- list(
  list(path = base_path_train, type = "Train"),
  list(path = base_path_test, type = "Test"),
  list(path = path_translation_train, type = "Train"),
  list(path = path_translation_test, type = "Test")
)
```

```{r}
# Function to determine language
get_language <- function(file_path, df) {
  if (str_detect(file_path, "mig-multi")) {
    return(df$lang[1])
  } else {
    filename <- basename(file_path)
    lang_from_name <- str_match(filename, "_([a-z]{2})\\.csv$|_([a-z]{2})_")[,2]
    if (!is.na(lang_from_name)) {
      return(lang_from_name)
    } else {
      return(df$tweet_lang[1])
    }
  }
}
```

```{r}
extract_model_dataset <- function(filename) {
  # Remove .csv and extract base name
  base <- basename(gsub("\\.csv$", "", filename))
  parts <- unlist(strsplit(base, "_"))

  # Try to extract model name: look for stop tokens like _K, _mig-*, _test
  stop_keywords <- grep("^K$|^mig-|^test$", parts)

  if (length(stop_keywords) > 0) {
    model <- paste(parts[1:(min(stop_keywords) - 1)], collapse = "_")
  } else {
    model <- base  # fallback: use whole filename if pattern unclear
  }

  # Convert list to named vector
  model_map <- c(
    "llama32-3B-mig-en-es-Q4" = "English-Spanish",
    "llama32-3B-mig-en-Q4"    = "English",
    "llama32-3B-mig-es-Q4"    = "Spanish",
    "llama32-3B-mig-multi-Q4" = "Multilanguage"
  )

  # Lookup in named vector
  if (model %in% names(model_map)) {
    return(model_map[model])
  } else {
    return(model)
  }
}

```


```{r}
lang_code_to_name <- function(code) {
  language_map <- list(
    "ar" = "Arabic",
    "es" = "Spanish",
    "pt" = "Portuguese",
    "fr" = "French",
    "it" = "Italian",
    "de" = "German",
    "hi" = "Hindi",
    "hu" = "Hungarian",
    "in" = "Indonesian",
    "tr" = "Turkish",
    "pl" = "Polish",
    "ko" = "Korean",
    "en" = "English"
  )
  return(ifelse(code %in% names(language_map), language_map[[code]], code))
}
```

```{r}
process_dataset <- function(file_path, split_type) {
  
  # Skip files that contain 'full' in the name
  if (str_detect(tolower(file_path), "full")) {
    #message(paste("Skipping file with 'full' in name:", file_path))
    return(NULL)
  }
  
  if (!grepl("Q4", basename(file_path))) {
    return(NULL)
  }
  
  # Read the file based on its extension (CSV or Excel)
  file_extension <- tools::file_ext(file_path)
  if (file_extension == "csv") {
    df <- read_csv(file_path, col_types = cols(.default = "c"))
  } else if (file_extension %in% c("xlsx", "xls")) {
    df <- read_excel(file_path)
  } else {
    return(NULL)  # Skip unsupported file types
  }
  
  # Rename "text" to "plaintext" only if "plaintext" doesn't already exist
  if ("text" %in% names(df) && !"plaintext" %in% names(df)) {
    df <- df %>% rename(plaintext = text)
  }
  
# Function to detect if a file is from mig-multi or one of the translation folders
is_mig_multi_or_translation <- function(file_path) {
  path <- tolower(normalizePath(file_path))
  translation_test <- tolower(normalizePath(path_translation_test))
  translation_train <- tolower(normalizePath(path_translation_train))

  str_detect(path, "mig-multi") |
    str_starts(path, translation_test) |
    str_starts(path, translation_train)
}
  
  # Determine if it's a relevant dataset
  mig_or_trans_flag <- is_mig_multi_or_translation(file_path)
  
  # Assign language column row-wise
  df <- df %>%
    mutate(lang_code = if (mig_or_trans_flag) lang else tweet_lang)

  df$Language <- map_chr(df$lang_code, lang_code_to_name)
  
  # assign message_id
  df <- df %>%
    mutate(
      message_id = case_when(
        is.numeric(message_id) ~ formatC(message_id, format = "f", drop0trailing = TRUE),
        TRUE ~ as.character(message_id)
      )
    )

  # Assign Model
  model <- extract_model_dataset(file_path)

  if ("translation_check" %in% colnames(df)) {
    df$TranslationQuality <- ifelse(df$translation_check == 1, "Good", "Bad")
    df$Translation <- 1
  } else if (startsWith(file_path, path_translation_train) || startsWith(file_path, path_translation_test)) {
    df$Translation <- 1
    df$TranslationQuality <- "Unknown"
  } else {
    df$Translation <- 0
    df$TranslationQuality <- "Not Translated"
  }

  # Final dataframe
  df_processed <- df %>%
    mutate(
      Model = model,
      Train_Test = split_type,
      Correct = ifelse(label_llama == label, 1, 0),
      file_path = file_path
    ) %>%
    select(
      message_id, Model, Language, Translation, TranslationQuality,
      Train_Test, Correct, label, label_llama, plaintext, file_path
    )
  
  # Rename Model values here
  df_processed <- df_processed %>%
    mutate(Model = case_when(
      Model == "llama-32-3B-en-es-Q4" ~ "English-Spanish",
      Model == "llama-32-3B-en-Q4"    ~ "English",
      Model == "llama-32-3B-es-Q4"    ~ "Spanish",
      Model == "llama-32-3B-multi-Q4" ~ "Multilanguage",
      TRUE ~ Model
    ))
  
  # make consistent Neutral
  df_processed <- df_processed %>%
    mutate(label = case_when(
      label == "Neutral" ~ "neutral",
      TRUE ~ label
    ))


  return(df_processed)
}
```

```{r}
# assign new IDs to each unique plaintext
assign_new_ids <- function(df) {
  df %>%
    group_by(plaintext) %>%
    mutate(newId = cur_group_id()) %>%
    ungroup()
}
```

### Create dataset
```{r}
# Collect all .csv files from test and train folders
test_files <- list.files(base_path_test, pattern = "\\.csv$", full.names = TRUE)
train_files <- list.files(base_path_train, pattern = "\\.csv$", full.names = TRUE)
translation_test_files <- list.files(path_translation_test, pattern = "\\.xlsx$", full.names = TRUE)
translation_train_files <- list.files(path_translation_train, pattern = "\\.xlsx$", full.names = TRUE)

# Process each file and combine
processed_test <- map_dfr(test_files, ~process_dataset(.x, "Test"))
processed_train <- map_dfr(train_files, ~process_dataset(.x, "Train"))
processed_translation_test <- map_dfr(translation_test_files, ~process_dataset(.x, "Test"))
processed_translation_train <- map_dfr(translation_train_files, ~process_dataset(.x, "Train"))

# Merge all data
df <- bind_rows(
  processed_train,
  processed_test,
  processed_translation_train,
  processed_translation_test
)
```

#### Post-processing dataset
```{r}
# assign new ID
df <- assign_new_ids(df)
```

```{r}
# count number of obs per language
df %>%
  count(Language, name = "n_observations") %>%
  arrange(desc(n_observations))

table(df$Language)
```

```{r}
# remove duplicate - each observation should have a unique pair of newId and filepath
df <- df %>%
  distinct(file_path, newId, .keep_all = TRUE) 
```

```{r}
#table(df2$Language, df2$Model)
```

```{r}
# remove obs where Language is NA
df <- df %>%
  filter(!is.na(Language)) %>%
  distinct(file_path, newId, .keep_all = TRUE)
```

```{r}
# remove all observation where label is NA
df <- df %>%
  filter(!is.na(label))
```

```{r}
# Ensure all variables are factors
df <- df %>%
  mutate(
    message_id = as.character(message_id),
    Model = as.factor(Model),
    Language = as.factor(Language),
    Train_Test = as.factor(Train_Test),
    Translation = as.factor(Translation),
    TranslationQuality = as.factor(TranslationQuality),
    Correct = as.factor(Correct)
  )
```

### Checks
```{r}
# check if there are NAs in df
na_count <- sapply(df, function(x) sum(is.na(x)))
print(na_count)

# check how many correct values are in Correct
table(df$Correct)
```

```{r}
# remove obs with NA in label_llama
df <- df %>%
  filter(!is.na(label_llama))
```

```{r}
df <- df %>%
  group_by(Language, Train_Test) %>%
  mutate(Share_Unrelated = mean(label == "unrelated")) %>%
  ungroup()
```


```{r}
# count number of obs per language
df %>%
  count(Language, name = "n_observations") %>%
  arrange(desc(n_observations))

table(df$Language)
```

```{r}
# count number of unique observations per language
df_unique_counts <- df %>%
  distinct(newId, Language) %>%  # ensure uniqueness of newID per language
  count(Language, name = "n_unique_observations") %>%
  arrange(desc(n_unique_observations))

df_unique_counts
```

```{r}
# count the number of obs with Translation == 'Good'
translation_count <- df %>%
  filter(TranslationQuality == 'Good') %>%
  group_by(Model, Language) %>%
  summarise(count = n())
translation_count
```

### Add language pre-trainign share from Touvron et al. (2023)
```{r}
# Create the language share lookup table
lang_share_lookup <- data.frame(
  code = c("en", "unknown", "de", "fr", "sv", "zh", "es", "ru", "nl", "it", "ja", "pl", "pt", "vi",
           "uk", "ko", "ca", "sr", "id", "cs", "fi", "hu", "no", "ro", "bg", "da", "sl", "hr"),
  Language = c("English", "Unknown", "German", "French", "Swedish", "Chinese", "Spanish", "Russian", "Dutch", "Italian", 
               "Japanese", "Polish", "Portuguese", "Vietnamese", "Ukrainian", "Korean", "Catalan", "Serbian", "Indonesian",
               "Czech", "Finnish", "Hungarian", "Norwegian", "Romanian", "Bulgarian", "Danish", "Slovene", "Croatian"),
  LangShare = c(89.70, 8.38, 0.17, 0.16, 0.15, 0.13, 0.13, 0.13, 0.12, 0.11, 0.10, 0.09, 0.09, 0.08,
                0.07, 0.06, 0.04, 0.04, 0.03, 0.03, 0.03, 0.03, 0.03, 0.03, 0.02, 0.02, 0.01, 0.01)
)

# Merge the LangShare into your existing df by the language code
df <- merge(df, lang_share_lookup, by = "Language", all.x = TRUE)
df$LangShare[is.na(df$LangShare)] <- 0
```

## Model
```{r}
# Set reference levels for the factors
df$Model <- relevel(df$Model, ref = "English") #llama-32-3B-en-Q4 - all the models are 4bit
df$Language <- relevel(df$Language, ref = "English")
df$Train_Test <- relevel(df$Train_Test, ref = "Train")
df$label <- factor(df$label)
df$label <- relevel(df$label, ref = "neutral")
df$TranslationQuality <- factor(df$TranslationQuality)
df$TranslationQuality <- relevel(df$TranslationQuality, ref = "Not Translated") # Not Translated
df$Translation <- factor(df$Translation)

# Fit the logistic regression model
logit_model <- glm(Correct ~ Model + Language + Train_Test + TranslationQuality + label, # add Translated not Translated and Translated(0,1)*TranslationQuality(unknown, bad, good)  
                   data = df, 
                   family = binomial())

# Summarize the model
summary(logit_model)
```

```{r}
# Model 2 - Interactions Model*Language
# Fit the logistic regression model
logit_modelxlanguage <- glm(Correct ~ Model*Language + label + Train_Test + TranslationQuality, # add Translated not Translated and Translated(0,1)*TranslationQuality(unknown, bad, good)  
                   data = df, 
                   family = binomial())

# Summarize the model
summary(logit_modelxlanguage)
```

```{r}
# Model 3 - Interactions Model*Language + Model*Share_Unrelated
# Fit the logistic regression model
logit_modelxShareUnrelated <- glm(Correct ~ Model*Language + Model*Share_Unrelated + Train_Test + TranslationQuality, # add Translated not Translated and Translated(0,1)*TranslationQuality(unknown, bad, good)
                   data = df, 
                   family = binomial())

# Summarize the model
summary(logit_modelxShareUnrelated)
```
###### Visualization of the interaction Model x Language + Model x Share of Unrelated
```{r}
# Define your languages and models
languages <- c("Turkish", "Spanish", "Portuguese", "Polish", "Korean", "Italian",
               "Indonesian", "Hungarian", "Hindi", "German", "French", "Arabic")
models <- c("English-Spanish", "Multilanguage", "Spanish")
shares <- seq(0, 1, 0.1)

# Extract coefficients from the model summary
coefs <- coef(logit_modelxShareUnrelated)
coef_names <- names(coefs)

# Function to get coefficient or return 0 if not found
get_coef <- function(term) {
  if (term %in% coef_names) {
    return(coefs[term])
  } else {
    return(0)
  }
}

# Build the dataset
df_plot <- expand.grid(
  Share_Unrelated = shares,
  Language = languages,
  Model = models
) %>%
  rowwise() %>%
  mutate(
    main_effect = get_coef(paste0("Model", Model)),
    lang_interaction = get_coef(paste0("Model", Model, ":Language", Language)),
    share_interaction = get_coef(paste0("Model", Model, ":Share_Unrelated")),
    # Replace NAs with 0
    #main_effect = replace_na(main_effect, 0),
    #lang_interaction = replace_na(lang_interaction, 0),
    #share_interaction = replace_na(share_interaction, 0),
    total_effect = main_effect + lang_interaction + Share_Unrelated * share_interaction,
    Accuracy = 1 / (1 + exp(-total_effect))  # logistic transformation
  )

# Plot
ggplot(df_plot, aes(x = Share_Unrelated, y = Accuracy, color = Model)) +
  geom_line(size = 1.1) +
  facet_wrap(~ Language, ncol = 4) +
  labs(
    x = "Share of Unrelated Tweets",
    y = "Accuracy", # or Log-Odds
    title = "Model Performance Across Languages and Noise Levels"
  ) +
  scale_x_continuous(labels = scales::percent_format(accuracy = 1)) +
  theme_minimal(base_size = 11) +
  theme(legend.position = "top")

# Save the plot
ggsave(
  filename = "predicted_accuracy_by_language.png",
  plot = last_plot(),
  path = 'output',
  width = 12,
  height = 8,
  dpi = 300
)
```


```{r}
# Model 4 - Interactions Model*label
# Fit the logistic regression model
logit_modelxlabel <- glm(Correct ~ Model*label + Language + Train_Test + TranslationQuality, # add Translated not Translated and Translated(0,1)*TranslationQuality(unknown, bad, good)  
                   data = df, 
                   family = binomial())

# Summarize the model
summary(logit_modelxlabel)
```

Idea - to test if the Multilanguage adds noise, we keep the language constant (English) and see how it performs. This means that we need to use only tweets that were classified based on the translation.

The results below seems to show that even when we use only tweets in English, the Multilanguage model performs worse than English-Spanish Model and Spanish Model in classifying unrelated tweets.
This seems to suggest that Multilanguage model is worse when it comes to classifying unrelated tweets, even when we use only tweets in English.

```{r}
# select only tweets classified based on the translation
df_translation <- df %>%
  filter(Translation == 1)

# Fit the logistic regression model
#logit_modelxlabel <- glm(Correct ~ Model*label + Language + Train_Test + TranslationQuality, # add Translated not Translated and Translated(0,1)*TranslationQuality(unknown, bad, good)  
#                   data = df_translation, 
#                   family = binomial())

# Summarize the model
#summary(logit_modelxlabel)
```


```{r}
# create log of LangShare
df$LogLangShare <- log1p(df$LangShare)

# Model 5 - Interactions Model and Language Share from LLAMA 2
logit_model_interact <- glm(Correct ~ Model*LangShare + Language + label + Train_Test + TranslationQuality + label,
                            data = df,
                            family = binomial())

summary(logit_model_interact)
```

```{r}
# convert Correct to numeric
df$Correct_Num <- as.numeric(as.character(df$Correct))

lang_summary <- df %>%
  group_by(Language, LangShare) %>%
  summarise(accuracy = mean(Correct_Num, na.rm = TRUE), n = n()) %>%
  ungroup()
```

```{r}
# Show relationship between accuracy and share of language in pretraining phase in LLAMA
# create summary table removing unrelated and English tweets
df_summary <- df %>%
  # remove obs where Language is English
  filter(Language != "English") %>%
  filter(label != "unrelated") %>%
  group_by(Language, Model, LangShare) %>%
  summarise(accuracy = mean(Correct_Num), .groups = "drop")

# plot accuracy by Language and Model
ggplot(df_summary, aes(x = LangShare, y = accuracy, color = Model)) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(title = "Accuracy by Language Share and Model Type",
       x = "Language Share in Pretraining (%)",
       y = "Mean Accuracy") +
  theme_minimal()

# save the plot in  folder
ggsave(file.path(output_dir, "accuracy_by_language_share.png"), width = 10, height = 6)
```

```{r}
# Step 1: Summarize and order Language by mean LangShare
language_order <- df_summary %>%
  group_by(Language) %>%
  summarise(mean_share = mean(LangShare), .groups = "drop") %>%
  arrange(mean_share) %>%
  pull(Language)
# Step 2: Set Language as ordered factor
df_summary$Language <- factor(df_summary$Language, levels = language_order)
```


```{r}
# Plot
ggplot(df_summary, aes(x = LangShare, y = accuracy, color = Language)) +
  geom_point(size = 2) +
  geom_smooth(method = "lm", se = FALSE, color = "black", aes(group = 1)) +
  facet_wrap(~ Model) +
  labs(
    title = "Accuracy vs Language Share by Model",
    x = "Share of Language in LLAMA Pretraining",
    y = "Mean Accuracy",
    color = "Language"
  ) +
  theme_minimal()

# save the plot in output folder
ggsave(file.path(output_dir, "accuracy_vs_language_share_by_model.png"), width = 10, height = 6)
```

```{r}
# Get linear model coefficients per model
coefficients_table <- df_summary %>%
  group_by(Model) %>%
  do(tidy(lm(accuracy ~ LangShare, data = .))) %>%
  filter(term == "LangShare") %>%
  select(Model, Estimate = estimate, Std_Error = std.error, p_value = p.value)

print(coefficients_table)
```


#### Multilevel Models
```{r}
library(lme4)

# Multilevel logistic regression
logit_mixed <- glmer(Correct ~ Model * Language + label + Train_Test + TranslationQuality + 
             (1 | Language), 
      data = df, family = binomial())

# Summarize the model
summary(logit_mixed)
```
```{r}
# Multilevel logistic regression with random slopes
multilevel_interaction <- glmer(Correct ~ Model + label + Train_Test + TranslationQuality + 
             (1 + Model | Language), 
      data = df, family = binomial())

summary(multilevel_interaction)

```

##### Random Forest
```{r}
tree <- rpart(Correct ~ Model + Language + Train_Test + TranslationQuality + label, # add TrainSize
                         data = df)
rpart.plot(tree, yesno = 2)
```
