---
title: "modelling_llm_performances"
author: "Andrea Nasuto"
date: "2025-04-08"
output: html_document
---

Build a model to assess the performance of a series of LLMs across multiple languages, translations, test/train dataset

The dataset to perform this should have this structure. All variables should be factor/characters:

ID | Model | Language | Translation | Test/Train | Correct
100100 | FT-Multi | IT | 0 | Train | 1
100100 | FT-EN | IT | 1 | Test | 0
100100 | FT-EN-ES | IT | 1 | Train | 1
120300 | LLM2 | EN | EN | Train | 1
...


```{r}
# Set the option to avoid scientific notation
options(scipen = 999)
```

```{r}
# load libraries
library(dplyr)
library(ggplot2)
library(tidyr)
library(purrr)
library(readr)
library(stringr)
library(broom)
library(caret)
library(readxl)
library(pROC)
library(rpart)
library(rpart.plot)
library(viridis)
library(showtext)
library(sysfonts)
library(forcats)
library(scales)
library(extrafont)
library(RColorBrewer)
library(stringr)

# Add Google Fonts (easier than importing system fonts)
font_add_google("Roboto Condensed", "roboto_condensed")
showtext_auto()
```

```{r}
# set working directory
knitr::opts_knit$set(root.dir = "/Users/Andrea/Library/CloudStorage/OneDrive-TheUniversityofLiverpool/PhD/code/immigration-llm")

# dataset paths
base_path_test <- "/Users/Andrea/Library/CloudStorage/OneDrive-TheUniversityofLiverpool/PhD/code/immigration-llm/model_performance/data/test_dataset_raw"

base_path_train <- "/Users/Andrea/Library/CloudStorage/OneDrive-TheUniversityofLiverpool/PhD/code/immigration-llm/model_performance/data/train_dataset_raw"

path_translation_test <- "/Users/Andrea/Library/CloudStorage/OneDrive-TheUniversityofLiverpool/PhD/code/immigration-llm/model_performance/data/test_dataset_translation"

path_translation_train <- "/Users/Andrea/Library/CloudStorage/OneDrive-TheUniversityofLiverpool/PhD/code/immigration-llm/model_performance/data/train_dataset_translation"

#path_translation_checked_test <- "/Users/Andrea/Library/CloudStorage/OneDrive-TheUniversityofLiverpool/PhD/code/global-immigration-sentiment/validation/tweet_translation_checked/test/"

#path_translation_checked_train <- "/Users/Andrea/Library/CloudStorage/OneDrive-TheUniversityofLiverpool/PhD/code/global-immigration-sentiment/validation/tweet_translation_checked/train/"

# output path
output_dir <- "output"
```

```{r}
# Get all CSV files
test_files <- list.files(base_path_test, pattern = "\\.csv$", full.names = TRUE)
train_files <- list.files(base_path_train, pattern = "\\.csv$", full.names = TRUE)

# Get all CSV files
test_translation_files <- list.files(path_translation_test, pattern = "\\.(csv|xlsx)$", full.names = TRUE)
train_translation_files <- list.files(path_translation_train, pattern = "\\.(csv|xlsx)$", full.names = TRUE)

# Define dataset directories
dirs <- list(
  list(path = base_path_train, type = "Train"),
  list(path = base_path_test, type = "Test"),
  list(path = path_translation_train, type = "Train"),
  list(path = path_translation_test, type = "Test")
)
```

```{r}
# Function to determine language
get_language <- function(file_path, df) {
  if (str_detect(file_path, "mig-multi")) {
    return(df$lang[1])
  } else {
    filename <- basename(file_path)
    lang_from_name <- str_match(filename, "_([a-z]{2})\\.csv$|_([a-z]{2})_")[,2]
    if (!is.na(lang_from_name)) {
      return(lang_from_name)
    } else {
      return(df$tweet_lang[1])
    }
  }
}
```

```{r}
extract_model_dataset <- function(filename) {
  # Remove .csv and extract base name
  base <- basename(gsub("\\.csv$", "", filename))
  parts <- unlist(strsplit(base, "_"))

  # Try to extract model name: look for stop tokens like _K, _mig-*, _test
  stop_keywords <- grep("^K$|^mig-|^test$", parts)

  if (length(stop_keywords) > 0) {
    model <- paste(parts[1:(min(stop_keywords) - 1)], collapse = "_")
  } else {
    model <- base  # fallback: use whole filename if pattern unclear
  }

  # Convert list to named vector
  model_map <- c(
    "llama32-3B-mig-en-es-Q4" = "English-Spanish",
    "llama32-3B-mig-en-Q4"    = "English",
    "llama32-3B-mig-es-Q4"    = "Spanish",
    "llama32-3B-mig-multi-Q4" = "Multilanguage"
  )

  # Lookup in named vector
  if (model %in% names(model_map)) {
    return(model_map[model])
  } else {
    return(model)
  }
}

```

```{r}
lang_code_to_name <- function(code) {
  language_map <- list(
    "ar" = "Arabic",
    "es" = "Spanish",
    "pt" = "Portuguese",
    "fr" = "French",
    "it" = "Italian",
    "de" = "German",
    "hi" = "Hindi",
    "hu" = "Hungarian",
    "in" = "Indonesian",
    "tr" = "Turkish",
    "pl" = "Polish",
    "ko" = "Korean",
    "en" = "English"
  )
  return(ifelse(code %in% names(language_map), language_map[[code]], code))
}
```

```{r}
process_dataset <- function(file_path, split_type) {
  
  # Skip files that contain 'full' in the name
  if (str_detect(tolower(file_path), "full")) {
    #message(paste("Skipping file with 'full' in name:", file_path))
    return(NULL)
  }
  
  if (!grepl("Q4", basename(file_path))) {
    return(NULL)
  }
  
  # Read the file based on its extension (CSV or Excel)
  file_extension <- tools::file_ext(file_path)
  if (file_extension == "csv") {
    df <- read_csv(file_path, col_types = cols(.default = "c"))
  } else if (file_extension %in% c("xlsx", "xls")) {
    df <- read_excel(file_path)
  } else {
    return(NULL)  # Skip unsupported file types
  }
  
  # Rename "text" to "plaintext" only if "plaintext" doesn't already exist
  if ("text" %in% names(df) && !"plaintext" %in% names(df)) {
    df <- df %>% rename(plaintext = text)
  }
  
# Function to detect if a file is from mig-multi or one of the translation folders
is_mig_multi_or_translation <- function(file_path) {
  path <- tolower(normalizePath(file_path))
  translation_test <- tolower(normalizePath(path_translation_test))
  translation_train <- tolower(normalizePath(path_translation_train))

  str_detect(path, "mig-multi") |
    str_starts(path, translation_test) |
    str_starts(path, translation_train)
}
  
  # Determine if it's a relevant dataset
  mig_or_trans_flag <- is_mig_multi_or_translation(file_path)
  
  # Assign language column row-wise
  df <- df %>%
    mutate(lang_code = if (mig_or_trans_flag) lang else tweet_lang)

  df$Language <- map_chr(df$lang_code, lang_code_to_name)
  
  # assign message_id
  df <- df %>%
    mutate(
      message_id = case_when(
        is.numeric(message_id) ~ formatC(message_id, format = "f", drop0trailing = TRUE),
        TRUE ~ as.character(message_id)
      )
    )

  # Assign Model
  model <- extract_model_dataset(file_path)

  if ("translation_check" %in% colnames(df)) {
    df$TranslationQuality <- ifelse(df$translation_check == 1, "Good", "Bad")
    df$Translation <- 1
  } else if (startsWith(file_path, path_translation_train) || startsWith(file_path, path_translation_test)) {
    df$Translation <- 1
    df$TranslationQuality <- "Unknown"
  } else {
    df$Translation <- 0
    df$TranslationQuality <- "Not Translated"
  }

  # Final dataframe
  df_processed <- df %>%
    mutate(
      Model = model,
      Train_Test = split_type,
      Correct = ifelse(label_llama == label, 1, 0),
      file_path = file_path
    ) %>%
    select(
      message_id, Model, Language, Translation, TranslationQuality,
      Train_Test, Correct, label, label_llama, plaintext, file_path
    )
  
  # Rename Model values here
  df_processed <- df_processed %>%
    mutate(Model = case_when(
      Model == "llama-32-3B-en-es-Q4" ~ "English-Spanish",
      Model == "llama-32-3B-en-Q4"    ~ "English",
      Model == "llama-32-3B-es-Q4"    ~ "Spanish",
      Model == "llama-32-3B-multi-Q4" ~ "Multilanguage",
      TRUE ~ Model
    ))
  
  # make consistent Neutral
  df_processed <- df_processed %>%
    mutate(label = case_when(
      label == "Neutral" ~ "neutral",
      TRUE ~ label
    ))


  return(df_processed)
}
```

```{r}
# assign new IDs to each unique plaintext
assign_new_ids <- function(df) {
  df %>%
    group_by(plaintext) %>%
    mutate(newId = cur_group_id()) %>%
    ungroup()
}
```

### Create dataset
```{r}
# Collect all .csv files from test and train folders
test_files <- list.files(base_path_test, pattern = "\\.csv$", full.names = TRUE)
train_files <- list.files(base_path_train, pattern = "\\.csv$", full.names = TRUE)
translation_test_files <- list.files(path_translation_test, pattern = "\\.xlsx$", full.names = TRUE)
translation_train_files <- list.files(path_translation_train, pattern = "\\.xlsx$", full.names = TRUE)

# Process each file and combine
processed_test <- map_dfr(test_files, ~process_dataset(.x, "Test"))
processed_train <- map_dfr(train_files, ~process_dataset(.x, "Train"))
processed_translation_test <- map_dfr(translation_test_files, ~process_dataset(.x, "Test"))
processed_translation_train <- map_dfr(translation_train_files, ~process_dataset(.x, "Train"))

# Merge all data
df <- bind_rows(
  processed_train,
  processed_test,
  processed_translation_train,
  processed_translation_test
)
```

#### Post-processing dataset
```{r}
# assign new ID
df <- assign_new_ids(df)
```

```{r}
# count number of obs per language
df %>%
  count(Language, name = "n_observations") %>%
  arrange(desc(n_observations))

table(df$Language)
```

```{r}
table(df$Language, df$Model)
```


```{r}
# Remove rows with NA in Language or plaintext
df <- df %>%
  filter(!is.na(Language)) %>%
  filter(!is.na(plaintext))
```

```{r}
df <- df %>%
  distinct(Model, newId, Translation, .keep_all = TRUE) # make sure that each Model has only one observation per newId and Translation
```

```{r}
# Ensure all variables are factors
df <- df %>%
  mutate(
    message_id = as.factor(message_id),
    Model = as.factor(Model),
    Language = as.factor(Language),
    Train_Test = as.factor(Train_Test),
    Translation = as.factor(Translation),
    TranslationQuality = as.factor(TranslationQuality),
    Correct = as.factor(Correct)
  )
```

```{r}
# show distribution of obs by Language and Model
table(df$Language, df$Model)
```

### Checks
```{r}
# check if there are NAs in df
na_count <- sapply(df, function(x) sum(is.na(x)))
print(na_count)

# check how many correct values are in Correct
table(df$Correct)
```

```{r}
# remove obs with NA in label_llama
df <- df %>%
  filter(!is.na(label_llama))
```

```{r}
df <- df %>%
  group_by(Language, Train_Test) %>%
  mutate(Share_Unrelated = mean(label == "unrelated")) %>%
  ungroup()
```


```{r}
# count number of obs per language
df %>%
  count(Language, name = "n_observations") %>%
  arrange(desc(n_observations))
```

```{r}
# count number of unique observations per language
df_unique_counts <- df %>%
  distinct(newId, Language) %>%  # ensure uniqueness of newID per language
  count(Language, name = "n_unique_observations") %>%
  arrange(desc(n_unique_observations))

df_unique_counts
```

```{r}
# count the number of obs with Translation == 'Good'
translation_count <- df %>%
  filter(TranslationQuality == 'Good') %>%
  group_by(Model, Language) %>%
  summarise(count = n())
translation_count
```

### Add language pre-trainign share from Touvron et al. (2023)
```{r}
# Create the language share lookup table
lang_share_lookup <- data.frame(
  code = c("en", "unknown", "de", "fr", "sv", "zh", "es", "ru", "nl", "it", "ja", "pl", "pt", "vi",
           "uk", "ko", "ca", "sr", "id", "cs", "fi", "hu", "no", "ro", "bg", "da", "sl", "hr"),
  Language = c("English", "Unknown", "German", "French", "Swedish", "Chinese", "Spanish", "Russian", "Dutch", "Italian", 
               "Japanese", "Polish", "Portuguese", "Vietnamese", "Ukrainian", "Korean", "Catalan", "Serbian", "Indonesian",
               "Czech", "Finnish", "Hungarian", "Norwegian", "Romanian", "Bulgarian", "Danish", "Slovene", "Croatian"),
  LangShare = c(89.70, 8.38, 0.17, 0.16, 0.15, 0.13, 0.13, 0.13, 0.12, 0.11, 0.10, 0.09, 0.09, 0.08,
                0.07, 0.06, 0.04, 0.04, 0.03, 0.03, 0.03, 0.03, 0.03, 0.03, 0.02, 0.02, 0.01, 0.01)
)

# Merge the LangShare into your existing df by the language code
df <- merge(df, lang_share_lookup, by = "Language", all.x = TRUE)
df$LangShare[is.na(df$LangShare)] <- 0
```

## Model
```{r}
# Set reference levels for the factors
df$Model <- relevel(df$Model, ref = "English") #llama-32-3B-en-Q4 - all the models are 4bit
df$Language <- relevel(df$Language, ref = "English")
df$Train_Test <- relevel(df$Train_Test, ref = "Train")
df$label <- factor(df$label)
df$label <- relevel(df$label, ref = "neutral")
df$TranslationQuality <- factor(df$TranslationQuality)
df$TranslationQuality <- relevel(df$TranslationQuality, ref = "Not Translated") # Not Translated
df$Translation <- factor(df$Translation)

# Fit the logistic regression model
logit_model <- glm(Correct ~ Model + Language + Train_Test + TranslationQuality + label, # add Translated not Translated and Translated(0,1)*TranslationQuality(unknown, bad, good)  
                   data = df, 
                   family = binomial())

# Summarize the model
summary(logit_model)
```

```{r}
# Load the font (only needs to be run once per session)
font_add_google("Roboto Condensed", "roboto")
showtext_auto()
```

```{r}
baseline_results <- tidy(logit_model, conf.int = TRUE) %>%
  filter(term != "(Intercept)") %>%
  mutate(
    variable_group = case_when(
      str_starts(term, "Model")              ~ "Model",
      str_starts(term, "Language")           ~ "Language",
      str_starts(term, "TranslationQuality") ~ "Translation",
      str_starts(term, "label")              ~ "Label",
      TRUE                                   ~ "Other"
    ),
    clean_term = case_when(
      variable_group == "Model"        ~ str_remove(term, "^Model"),
      variable_group == "Language"     ~ str_remove(term, "^Language"),
      variable_group == "Translation"  ~ str_remove(term, "^TranslationQuality"),
      variable_group == "Label"        ~ case_when(
        str_detect(term, "unrelated") ~ "Unrelated",
        str_detect(term, "anti")      ~ "Anti-Immigration",
        str_detect(term, "pro")       ~ "Pro-Immigration",
        TRUE                          ~ str_remove(term, "^label")
      ),
      TRUE                             ~ term
    ),
    significance = case_when(
      p.value < 0.001 ~ "***",
      p.value < 0.01  ~ "**", 
      p.value < 0.05  ~ "*",
      TRUE            ~ ""
    ),
    abs_estimate = abs(estimate)
  ) %>%
  # Remove Train_Test effects
  filter(!str_detect(term, "^Train_Test")) %>%
  arrange(variable_group, desc(abs_estimate))
```

```{r}
# set color blind palette
cb_palette <- c(
  "#E69F00", "#56B4E9", "#009E73", "#F0E442",
  "#0072B2", "#D55E00", "#CC79A7", "#999999"
)
```

```{r}
# Add Roboto Condensed
font_add_google("Roboto Condensed", "roboto_condensed")

# Turn on showtext
showtext_auto()
```

```{r}
# Define named palette based on cb_palette positions
cb_palette_named <- c(
  "Model"       = cb_palette[1],
  "Language"    = cb_palette[5],
  "Label"       = cb_palette[3],
  "Translation" = cb_palette[6]
)

baseline_plot <- baseline_results %>%
  mutate(
    clean_term = fct_reorder(clean_term, estimate),
    variable_group = fct_reorder(variable_group, abs_estimate, .fun = median)
  ) %>%
  ggplot(aes(x = estimate, y = clean_term, color = variable_group)) +
  geom_vline(xintercept = 0, linetype = "dashed", alpha = 0.7, linewidth = 0.5) +
  geom_errorbarh(aes(xmin = conf.low, xmax = conf.high),
                 height = 0.4, alpha = 1, linewidth = 0.6) +
  geom_point(size = 2.0, alpha = 1) +
  facet_grid(rows = vars(variable_group), scales = "free_y", space = "free_y", switch = "y") +
  scale_color_manual(values = cb_palette_named, guide = "none") +
  labs(
    #title = "Baseline Model",
    x = "Coefficient Estimates", y = NULL
  ) +
  theme(
    text              = element_text(family = "roboto_condensed"),
    strip.text.y.left = element_text(size = 11, face = "bold"),
    axis.text.y       = element_text(size = 12, lineheight = 0.8),
    axis.text.x       = element_text(size = 10),
    axis.title.x    = element_text(size = 14, margin = margin(t = 12)),
    panel.grid.minor  = element_blank(),
    strip.placement   = "outside",
    plot.margin       = margin(10, 30, 10, 10)  # give more room on the right
    #theme(
    #plot.title = element_text(size = 16, face = "bold"),
    #plot.subtitle = element_text(size = 12, color = "grey40"),
    #strip.text = element_text(size = 11, face = "bold"),
    #axis.text.y = element_text(size = 13),
    #axis.text.x = element_text(size = 10),
    #axis.title.x    = element_text(size = 13),
    #panel.grid.minor = element_blank()
    
  )

baseline_plot

# save the plot
ggsave(filename = "model1_baseline_plot.png", plot = baseline_plot, path = "output", width = 10, height = 6, dpi = 300)
```


#### Model 2
```{r}
# Model 2 - Interactions Model*Language
# Fit the logistic regression model
logit_modelxlanguage <- glm(Correct ~ Model*Language + label + Train_Test + TranslationQuality, # add Translated not Translated and Translated(0,1)*TranslationQuality(unknown, bad, good)  
                   data = df, 
                   family = binomial())

# Summarize the model
summary(logit_modelxlanguage)
```

```{r}
# Tidy and filter model 2 results
model2_results <- tidy(logit_modelxlanguage, conf.int = TRUE) %>%
  filter(term != "(Intercept)") %>%
  filter(
    grepl("^Model", term) |
    grepl("^Language", term)
  ) %>%
  mutate(
    variable_group = case_when(
      grepl("^Model:Language", term) ~ "Model × Language",
      grepl("^Model", term) ~ "Model Type",
      grepl("^Language", term) ~ "Language",
      TRUE ~ "Other"
    ),
    clean_term = case_when(
      grepl("^Model:Language", term) ~ {
        m <- gsub("^Model", "", term)
        m <- gsub(":Language", " × ", m)
        m
      },
      grepl("^Model", term) ~ gsub("^Model", "", term),
      grepl("^Language", term) ~ gsub("^Language", "", term),
      TRUE ~ tools::toTitleCase(term)
    ),
    abs_estimate = abs(estimate)
  )

# Plot 2 for Model 2
model2_plot2 <- model2_results %>%
  mutate(
    clean_term = fct_reorder(clean_term, estimate),
    variable_group = fct_reorder(variable_group, abs_estimate, .fun = median)
  ) %>%
  ggplot(aes(x = estimate, y = clean_term, color = variable_group)) +
  geom_vline(xintercept = 0, linetype = "dashed", alpha = 0.7, linewidth = 0.5) +
  geom_errorbarh(aes(xmin = conf.low, xmax = conf.high), height = 0.2, alpha = 0.7, linewidth = 0.6) +
  geom_point(size = 1.5, alpha = 1) +
  facet_wrap(~variable_group, scales = "free_y", ncol = 1, strip.position = "left") +
  scale_color_manual(values = cb_palette, guide = "none") +
  labs(
    #title = "Model × Language Interaction Coefficients",
    x = "Log-odds (Coefficient Estimate)", y = NULL
  ) +
  theme(
    text = element_text(family = "roboto_condensed"),
    plot.title = element_text(size = 16, face = "bold"),
    plot.subtitle = element_text(size = 12, color = "grey40"),
    strip.text = element_text(size = 9, face = "bold"),
    axis.text.y = element_text(size = 4),
    axis.text.x = element_text(size = 10),
    panel.grid.minor = element_blank()
  )

# show the plot
model2_plot2

# Save plot (uncomment to activate)
#ggsave(filename = "model2_plot2.png", plot = model2_plot2, path = "output", width = 10, height = 6, dpi = 300)
```


```{r}
# Tidy model
model2_tidy <- tidy(logit_modelxlanguage, conf.int = TRUE) %>%
  filter(term != "(Intercept)")

# Get model and language levels
model_levels <- levels(df$Model)
language_levels <- levels(df$Language)

# Create all combinations
combinations <- expand.grid(Model = model_levels, Language = language_levels, stringsAsFactors = FALSE)

# Function to get coefficient value or 0 if not present
get_coef <- function(term_name, coef_df) {
  val <- coef_df$estimate[coef_df$term == term_name]
  if (length(val) == 0) return(0)
  return(val)
}

# Sum main effects + interaction effects
combinations$log_odds <- mapply(function(model, lang) {
  main_model <- paste0("Model", model)
  main_lang  <- paste0("Language", lang)
  interaction <- paste0("Model", model, ":Language", lang)

  get_coef(main_model, model2_tidy) +
    get_coef(main_lang, model2_tidy) +
    get_coef(interaction, model2_tidy)
}, combinations$Model, combinations$Language)

# Add predicted accuracy
combinations$accuracy <- plogis(combinations$log_odds)  # inverse logit

```

```{r}
combinations$Model <- factor(
  combinations$Model,
  levels = c("English", "Spanish", "English-Spanish", "Multilanguage")
)

# 2. Compute mean accuracy by language and reorder Language (y-axis)
accuracy_grouped_combination <- combinations %>%
  group_by(Language) %>%
  summarise(mean_accuracy = mean(accuracy, na.rm = TRUE)) %>%
  arrange(mean_accuracy)  # most accurate at top

combinations$Language <- factor(
  combinations$Language,
  levels = accuracy_grouped_combination$Language
)
```

```{r}
# Heatmap of log-odds
heatmap_logodds <- ggplot(combinations, aes(x = Model, y = Language, fill = log_odds)) +
  geom_tile(color = "white") +
  scale_fill_viridis_c(
    option = "viridis",
    direction = 1,
    name = "Log-odds"
  ) +
  #labs(title = "Log-odds by Model × Language") +
  theme_minimal() +
  theme(
    text = element_text(family = "roboto_condensed"),
    plot.title = element_text(family = "roboto_condensed", face = "bold"),
    axis.text = element_text(family = "roboto_condensed"),
    legend.title = element_text(family = "roboto_condensed"),
    legend.text = element_text(family = "roboto_condensed"),
    axis.text.y = element_text(size = 11),
    axis.text.x = element_text(size = 11),
    axis.title.x    = element_text(size = 13, margin = margin(t = 12)),
    axis.title.y    = element_text(size = 13)
    )

# Heatmap of predicted accuracy
heatmap_accuracy <- ggplot(combinations, aes(x = Model, y = Language, fill = accuracy)) +
  geom_tile(color = "white") +
  scale_fill_gradient(low = "white", high = "darkgreen", name = "Predicted Accuracy") +
  #labs(title = "Predicted Accuracy by Model × Language") +
  theme_minimal() +
  theme(
    text = element_text(family = "roboto_condensed"),
    plot.title = element_text(family = "roboto_condensed", face = "bold"),
    axis.text = element_text(family = "roboto_condensed"),
    legend.title = element_text(family = "roboto_condensed"),
    legend.text = element_text(family = "roboto_condensed")
    #panel.grid.minor = element_blank()
  )

# Save the heatmaps
 ggsave(filename = "model_language_interaction_heatmap_logodds.png", plot = heatmap_logodds, path = "output", width = 8, height = 6, dpi = 300)
 #ggsave(filename = "model_language_interaction_heatmap_accuracy.png", plot = heatmap_accuracy, path = "output", width = 8, height = 6, dpi = 300)

# Print plots
print(heatmap_logodds)
#print(heatmap_accuracy)
```

#### Model 3
```{r}
# Model 3 - Interactions Model*Language + Model*Share_Unrelated
# Fit the logistic regression model
logit_modelxShareUnrelated <- glm(Correct ~ Model*Language + Model*Share_Unrelated + Train_Test + TranslationQuality, # add Translated not Translated and Translated(0,1)*TranslationQuality(unknown, bad, good)
                   data = df, 
                   family = binomial())

# Summarize the model
summary(logit_modelxShareUnrelated)
```

```{r}
# Color-blind–safe colours (Okabe–Ito + grey)
model_cols <- c(
  "English"         = "#9A9A9A",  # neutral mid-grey
  "Spanish"         = "#E69F00",  # orange from Okabe–Ito
  "English-Spanish" = "#FE6100",  # reddish-purple
  "Multilanguage"   = "#0072B2"   # deep blue
)

p <- ggplot(df_plot_model3, aes(x = Share_Unrelated, y = Accuracy, color = Model)) +
  geom_line(size = 1.2) +
  facet_wrap(~ Language, ncol = 4) +
  scale_color_manual(values = model_cols) +
  labs(
    x     = "Share of Unrelated Tweets",
    y     = "Mean Accuracy",
    color = "Model"
  ) +
  scale_x_continuous(labels = scales::percent_format(accuracy = 1)) +
  theme_minimal(base_size = 16) +
  theme(
    text             = element_text(family = "roboto_condensed"),
    axis.title       = element_text(size = 18),
    axis.text        = element_text(size = 14),
    strip.text       = element_text(size = 16),
    legend.title     = element_text(size = 16),
    legend.text      = element_text(size = 14),
    legend.position  = "top",
    panel.grid.minor = element_blank()
  )

#save plot
ggsave(filename = "model_shareunrelated_language_interaction.png", plot = p, path = "output", width = 12, height = 8, dpi = 300)
#print(p)
```
The chart if using the interaction between Language and Model (Model*Language), the results change based on the fact that for certain languages the Multilanguage model is better than others when the share of unrelated tweets is quite high (eg. 80%).

On the other hand, if the chart is using JUST the interaction between Model and Share_Unrelated (Model*Share_Unrelated), the results are more consistent across languages, showing that the Multilanguage model is generally worse than the English-Spanish model and Spanish model after 50% of tweets are not related to immigration. This means that the behavior is more standard.

```{r}
# Tidy and filter model 3 results
model3_results <- tidy(logit_modelxShareUnrelated, conf.int = TRUE) %>%
  filter(term != "(Intercept)") %>%
  filter(
    grepl("^Model", term) |
    grepl("Share_Unrelated", term)
  ) %>%
  mutate(
    variable_group = case_when(
      grepl("^Model:", term) ~ "Model × Share Unrelated",
      grepl("^Model", term) ~ "Model Type",
      grepl("Share_Unrelated", term) ~ "Share Unrelated",
      TRUE ~ "Other"
    ),
    clean_term = case_when(
      grepl("^Model:.*Share_Unrelated", term) ~ gsub("Model", "", term) %>%
        gsub(":Share_Unrelated", " × Share Unrelated", .) %>%
        gsub("_", " ", .),  # Replace underscores with spaces
      grepl("^Model", term) ~ gsub("^Model", "", term) %>%
        gsub("_", " ", .),  # Replace underscores with spaces
      TRUE ~ tools::toTitleCase(gsub("_", " ", term))  # Replace underscores with spaces
    ),
    abs_estimate = abs(estimate)
  )

# Plot 2 for Model 3
model3_plot2 <- model3_results %>%
  mutate(
    clean_term = fct_reorder(clean_term, estimate),
    variable_group = fct_reorder(variable_group, abs_estimate, .fun = median)
  ) %>%
  ggplot(aes(x = estimate, y = clean_term, color = variable_group)) +
  geom_vline(xintercept = 0, linetype = "dashed", alpha = 0.7, linewidth = 0.5) +
  geom_errorbarh(aes(xmin = conf.low, xmax = conf.high), height = 0.2, alpha = 0.7, linewidth = 0.6) +
  geom_point(size = 2.5, alpha = 0.9) +
  facet_wrap(~variable_group, scales = "free_y", ncol = 1, strip.position = "left") +
  scale_color_manual(values = cb_palette, guide = "none") +
  labs(
    #title = "Model × Share_Unrelated Interaction Coefficients",
    x = "Estimates", y = NULL
  ) +
  theme(
    text = element_text(family = "roboto_condensed"),
    plot.title = element_text(size = 16, face = "bold"),
    plot.subtitle = element_text(size = 12, color = "grey40"),
    strip.text = element_text(size = 9, face = "bold"),
    axis.text.y = element_text(size = 9),
    axis.text.x = element_text(size = 10),
    panel.grid.minor = element_blank()
  )

model3_plot2

# Save plot
ggsave(filename = "model_interaction_language_share_unrelated_plot.png", plot = model3_plot2, path = "output", width = 10, height = 6, dpi = 300)
```


#### Model 4


Idea - to test if the Multilanguage adds noise, we keep the language constant (English) and see how it performs. This means that we need to use only tweets that were classified based on the translation.

The results below seems to show that even when we use only tweets in English, the Multilanguage model performs worse than English-Spanish Model and Spanish Model in classifying unrelated tweets.
This seems to suggest that Multilanguage model is worse when it comes to classifying unrelated tweets, even when we use only tweets in English.

```{r}
model4_results <- tidy(logit_modelxlabel, conf.int = TRUE) %>%
  filter(term != "(Intercept)",
         term != "Train_TestTest",
         !grepl("TranslationQuality", term),
         !grepl("^Language", term)) %>%
  mutate(
    variable_group = case_when(
      grepl("^Model", term) ~ "Model Type",
      grepl("^label", term) ~ "Label Category", 
      grepl("Train_Test", term) ~ "Data Split",
      TRUE ~ "Other"
    ),
    # Extract components for clean label
    model_label = case_when(
      grepl("English-Spanish", term) ~ "English-Spanish",
      grepl("Multilanguage", term) ~ "Multilingual",
      grepl("Spanish", term) & variable_group == "Model Type" ~ "Spanish-Only",
      TRUE ~ NA_character_
    ),
    label_class = case_when(
      grepl("unrelated", term) ~ "Unrelated",
      grepl("anti", term) ~ "Anti-Immigration",
      grepl("pro", term) ~ "Pro-Immigration",
      TRUE ~ NA_character_
    ),
    # Compose readable label (use × for clarity)
    clean_term = case_when(
      !is.na(model_label) & !is.na(label_class) ~ paste(model_label, "×", label_class),
      !is.na(label_class) ~ label_class,
      !is.na(model_label) ~ model_label,
      TRUE ~ tools::toTitleCase(gsub("^Model|^label|^Language|^Train_Test|^TranslationQuality", "", term))
    ),
    significance = case_when(
      p.value < 0.001 ~ "***",
      p.value < 0.01 ~ "**", 
      p.value < 0.05 ~ "*",
      TRUE ~ ""
    ),
    abs_estimate = abs(estimate)
  ) %>%
  arrange(variable_group, desc(abs_estimate))
```

```{r}
model4_plot1 <- model4_results %>%
  mutate(clean_term = fct_reorder(clean_term, estimate)) %>%
  ggplot(aes(x = estimate, y = clean_term, color = variable_group)) +
  geom_vline(xintercept = 0, linetype = "dashed", alpha = 0.7, linewidth = 0.5) +
  geom_errorbarh(aes(xmin = conf.low, xmax = conf.high), height = 0.3, alpha = 0.8, linewidth = 0.8) +
  geom_point(size = 3, alpha = 0.9) +
  scale_color_manual(values = cb_palette, name = "Variable Group") +
  labs(
    #title = "Logistic Regression Coefficients",
    subtitle = "Log-odds estimates with 95% confidence intervals",
    x = "Coefficient Estimates", y = NULL,
    caption = "Dashed line at zero indicates no effect"
  ) +
  #theme_minimal(base_family = "Roboto Condensed") +
  theme(
    #plot.title = element_text(size = 16, face = "bold"),
    plot.subtitle = element_text(size = 12, color = "grey40"),
    axis.text.y = element_text(size = 9),
    axis.text.x = element_text(size = 10),
    axis.title.x = element_text(size = 11),
    legend.position = "bottom",
    legend.title = element_text(size = 10),
    panel.grid.minor = element_blank()
  )
```

```{r}
model4_plot2 <- model4_results %>%
  mutate(
    clean_term = fct_reorder(clean_term, estimate),
    variable_group = fct_reorder(variable_group, abs_estimate, .fun = median)
  ) %>%
  ggplot(aes(x = estimate, y = clean_term, color = variable_group)) +
  geom_vline(xintercept = 0, linetype = "dashed", alpha = 0.7, linewidth = 0.5) +
  geom_errorbarh(aes(xmin = conf.low, xmax = conf.high), height = 0.2, alpha = 0.7, linewidth = 0.6) +
  geom_point(size = 2.5, alpha = 0.9) +
  facet_wrap(~variable_group, scales = "free_y", ncol = 1, strip.position = "left") +
  scale_color_manual(values = cb_palette, guide = "none") +
  labs(
    #title = "LLM x Label Interaction Coefficients",
    x = "Coefficient Estimates", y = NULL
  ) +
  theme_minimal(base_family = "roboto_condensed") +
  theme(
    #plot.title = element_text(size = 16, face = "bold"),
    #plot.subtitle = element_text(size = 12, color = "grey40"),
    strip.text = element_text(size = 11, face = "bold"),
    axis.text.y = element_text(size = 13),
    axis.text.x = element_text(size = 10),
    axis.title.x    = element_text(size = 14, margin = margin(t = 12)),
    panel.grid.minor = element_blank()
    
  )

# save the plot
ggsave(filename = "model_label_interaction.png", plot = model4_plot2, path = "output", width = 10, height = 6, dpi = 300)
```

```{r}
model4_plot3 <- model4_results %>%
  mutate(
    clean_term = fct_reorder(clean_term, estimate),
    sig_shape = case_when(
      p.value < 0.001 ~ 19,
      p.value < 0.01 ~ 17,
      p.value < 0.05 ~ 15,
      TRUE ~ 1
    )
  ) %>%
  ggplot(aes(x = estimate, y = clean_term)) +
  geom_vline(xintercept = 0, linetype = "dashed", alpha = 0.7, linewidth = 0.5) +
  geom_segment(aes(x = conf.low, xend = conf.high, y = clean_term, yend = clean_term,
                   color = variable_group), alpha = 0.6, linewidth = 1) +
  geom_point(aes(color = variable_group, shape = factor(sig_shape)), size = 3) +
  scale_color_manual(values = cb_palette, name = "Variable Group") +
  scale_shape_manual(values = c(1, 15, 17, 19),
                     name = "Significance",
                     labels = c("n.s.", "p < 0.05", "p < 0.01", "p < 0.001")) +
  labs(
    #title = "Logistic Regression Coefficients with Significance Levels",
    x = "Log-odds (Coefficient Estimate)", y = NULL
  ) +
  #theme_minimal(base_family = "Roboto Condensed") +
  theme(
    #plot.title = element_text(size = 16, face = "bold"),
    #plot.subtitle = element_text(size = 12, color = "grey40"),
    axis.text.y = element_text(size = 9),
    axis.text.x = element_text(size = 10),
    legend.position = "bottom",
    panel.grid.minor = element_blank()
  )
```

```{r}
model4_plot4 <- model4_results %>%
  filter(!variable_group %in% c("Translation Quality", "Language")) %>%
  mutate(readable_label = clean_term) %>%
  mutate(readable_label = fct_reorder(readable_label, estimate)) %>%
  ggplot(aes(y = readable_label, color = variable_group, fill = variable_group)) +
  geom_vline(xintercept = 0, linetype = "solid", alpha = 0.3, linewidth = 1) +
  geom_col(aes(x = estimate), alpha = 0.6, width = 0.6) +
  geom_errorbarh(aes(xmin = conf.low, xmax = conf.high), 
                 height = 0.2, alpha = 0.8, linewidth = 0.8, color = "black") +
  geom_point(aes(x = estimate), size = 1.5, color = "black") +
  scale_fill_manual(values = cb_palette, name = "Variable Type") +
  scale_color_manual(values = cb_palette, name = "Variable Type") +
  labs(
   # title = "Model Performance: Share of Unrelated Classifications",
    x = "Coefficient Estimates", y = NULL
  ) +
  theme_minimal(base_family = "roboto_condensed") +
  theme(
    #plot.title = element_text(size = 16, face = "bold"),
    #plot.subtitle = element_text(size = 12, color = "grey40"),
    plot.caption = element_text(size = 9, color = "grey50", hjust = 0),
    axis.text.y = element_text(size = 9, color = "grey20"),
    axis.text.x = element_text(size = 10),
    legend.position = "bottom",
    panel.grid.minor = element_blank()
  )
```

```{r}
# print plots from model 4
#print(model4_plot1)
#print(model4_plot2)
#print(model4_plot3)
#print(model4_plot4)

# save plots from model 4
#mapply(function(plot, name) ggsave(filename = paste0(name, ".png"), plot = plot, path = "output", width = 10, height = 6, dpi = 300),
#       list(model_label_plot1, model_label_plot1, model_label_plot1, model_label_plot1),
#       paste0("model4_plot", 1:4))
```


Create heatmaps of Model 4 results across labels ana languages

```{r}
# Define models, languages, and labels
languages <- c("Turkish", "Spanish", "Portuguese", "Polish", "Korean", "Italian",
               "Indonesian", "Hungarian", "Hindi", "German", "French", "Arabic")
models <- c("English-Spanish", "Multilanguage", "Spanish", "English")
labels <- c("Neutral", "Anti-immigration", "Pro-immigration", "Unrelated")

# Set translation quality to a fixed level for prediction
translation_val <- "Unknown"  # or use "Good", "Bad"

# Get coefficients
coefs <- coef(logit_modelxlabel)
coef_names <- names(coefs)

# Function to extract coefficient or return 0
get_coef <- function(term) {
  if (term %in% coef_names) return(coefs[term]) else return(0)
}

# Average language effect (approximate for 'LanguageEnglish')
get_lang_coef <- function(language) {
  term <- paste0("Language", language)
  return(get_coef(term))
}

# Function to retrieve translation coefficient
get_translation_coef <- function(quality) {
  term <- paste0("TranslationQuality", quality)
  return(get_coef(term))
}

# Main label effect
get_label_coef <- function(label) {
  if (label == "Neutral") return(0)
  return(get_coef(paste0("label", gsub("-", "", label))))
}

# Model × Label interaction
get_model_label_interaction <- function(model, label) {
  term <- paste0("Model", model, ":label", gsub("-", "", label))
  return(get_coef(term))
}

# Build full table of predictions
df_plot_heatmap <- expand.grid(Language = languages, Model = models, Label = labels) %>%
  rowwise() %>%
  mutate(
    intercept = get_coef("(Intercept)"),
    model_coef = get_coef(paste0("Model", Model)),
    label_coef = get_label_coef(Label),
    model_label_interaction = get_model_label_interaction(Model, Label),
    language_coef = get_lang_coef(Language),
    test_coef = get_coef("Train_TestTest"),
    translation_coef = get_translation_coef(translation_val),
    total_eta = intercept + model_coef + label_coef + model_label_interaction + language_coef + translation_coef + test_coef,
    Accuracy = 1 / (1 + exp(-total_eta))
  ) %>%
  ungroup()

# Plot
ggplot(df_plot_heatmap, aes(x = Model, y = Language, fill = Accuracy)) +
  geom_tile(color = "white") +
  geom_text(
    aes(label = sprintf("%.1f", Accuracy * 100)),
    size = 3.2,
    family = "roboto",
    color = "white" # change color of the text
  ) +
  facet_wrap(~Label, ncol = 2) +
  scale_fill_viridis(
    name = "Accuracy",
    option = "D",
    labels = scales::percent_format(accuracy = 1)
  ) +
  labs(
    title = "Model 4 Accuracy Across Labels and Languages",
    x = NULL,
    y = NULL
  ) +
  #theme_minimal(base_size = 12, base_family = "roboto") +
  theme(
    legend.position = "right",
    strip.text = element_text(face = "bold"),
    axis.text.x = element_text(angle = 45, hjust = 1),
    axis.text.y = element_text(size = 10),
    axis.title = element_blank(),
    plot.title = element_text(face = "bold", size = 14, margin = margin(b = 10))
  )

# Save the plot
#ggsave(filename = "predicted_accuracy_by_labels_language.png", plot = last_plot(), path = 'output', width = 12, height = 8,dpi = 300)
```

Translated only observation analysis
```{r}
# select only tweets classified based on the translation
df_translation <- df %>%
  filter(Translation == 1)

# Fit the logistic regression model
#logit_modelxlabel <- glm(Correct ~ Model*label + Language + Train_Test + TranslationQuality, # add Translated not Translated and Translated(0,1)*TranslationQuality(unknown, bad, good)  
#                   data = df_translation, 
#                   family = binomial())

# Summarize the model
#summary(logit_modelxlabel)
```

Understand how pretraining impacts accuracy performances
```{r}
# create log of LangShare
df$LogLangShare <- log1p(df$LangShare)

# Model 5 - Interactions Model and Language Share from LLAMA 2
logit_model_interact <- glm(Correct ~ Model*LangShare + Language + label + Train_Test + TranslationQuality + label,
                            data = df,
                            family = binomial())

summary(logit_model_interact)
```

```{r}
# convert Correct to numeric
df$Correct_Num <- as.numeric(as.character(df$Correct))

lang_summary <- df %>%
  group_by(Language, LangShare) %>%
  summarise(accuracy = mean(Correct_Num, na.rm = TRUE), n = n()) %>%
  ungroup()
```

```{r}
# Show relationship between accuracy and share of language in pretraining phase in LLAMA
# create summary table removing unrelated and English tweets
df_summary <- df %>%
  # remove obs where Language is English
  filter(Language != "English") %>%
  filter(label != "unrelated") %>%
  group_by(Language, Model, LangShare) %>%
  summarise(accuracy = mean(Correct_Num), .groups = "drop")

# plot accuracy by Language and Model
ggplot(df_summary, aes(x = LangShare, y = accuracy, color = Model)) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(title = "Accuracy by Language Share and Model Type",
       x = "Language Share in Pretraining (%)",
       y = "Mean Accuracy") +
  theme_minimal()

# save the plot in  folder
#ggsave(file.path(output_dir, "accuracy_by_language_share.png"), width = 10, height = 6)
```

```{r}
# Step 1: Summarize and order Language by mean LangShare
language_order <- df_summary %>%
  group_by(Language) %>%
  summarise(mean_share = mean(LangShare), .groups = "drop") %>%
  arrange(mean_share) %>%
  pull(Language)
# Step 2: Set Language as ordered factor
df_summary$Language <- factor(df_summary$Language, levels = language_order)
```


```{r}
# Plot
ggplot(df_summary, aes(x = LangShare, y = accuracy, color = Language)) +
  geom_point(size = 2) +
  geom_smooth(method = "lm", se = FALSE, color = "black", aes(group = 1)) +
  facet_wrap(~ Model) +
  labs(
    #title = "Accuracy vs Language Share by Model",
    x = "Share of Language in LLAMA Pretraining",
    y = "Mean Accuracy",
    color = "Language"
  ) +
  theme_minimal(base_size = 16) +
  theme(
    text = element_text(family = "roboto_condensed"),
    axis.title = element_text(size = 18),
    axis.text = element_text(size = 14),
    legend.title = element_text(size = 16),
    legend.text = element_text(size = 14),
    strip.text = element_text(size = 14),
    axis.title.x = element_text(margin = margin(t = 15))
  )
# save the plot in output folder
ggsave(file.path(output_dir, "accuracy_vs_language_share_by_model.png"), width = 10, height = 6)
```

```{r}
# Get linear model coefficients per model
coefficients_table <- df_summary %>%
  group_by(Model) %>%
  do(tidy(lm(accuracy ~ LangShare, data = .))) %>%
  filter(term == "LangShare") %>%
  select(Model, Estimate = estimate, Std_Error = std.error, p_value = p.value)

print(coefficients_table)
```

#### Confusion Matrix and Confidence Intervals
```{r}
# Filter out translation data and separate train/test sets
df_non_translation <- df %>% filter(Translation == 0)

df_test <- df_non_translation %>% filter(Train_Test == "Test")
df_train <- df_non_translation %>% filter(Train_Test == "Train")
```

```{r}
# Function to calculate accuracy and 95% CI
compute_accuracy_ci <- function(data) {
  cm <- confusionMatrix(as.factor(data$label_llama), as.factor(data$label))
  accuracy <- cm$overall["Accuracy"]
  ci <- prop.test(cm$table %>% diag() %>% sum(), sum(cm$table))$conf.int
  return(tibble(
    Accuracy = accuracy,
    Accuracy_Lower = ci[1],
    Accuracy_Upper = ci[2]
  ))
}
```

```{r}
# Helper function to apply to each dataset split
generate_accuracy_df <- function(df_split, dataset_name) {
  df_split %>%
    group_by(Model, Language) %>%
    group_map(~{
      acc <- compute_accuracy_ci(.x)
      acc$Model <- .y$Model
      acc$Language <- .y$Language
      acc$Dataset <- dataset_name
      acc
    }) %>%
    bind_rows()
}
```

```{r}
generate_and_save_cm <- function(data, title) {
  models <- unique(data$Model)
  results <- list()

  for (model in models) {
    model_data <- data[data$Model == model, ]

    # Compute overall confusion matrix
    cm <- confusionMatrix(as.factor(model_data$label_llama), as.factor(model_data$label))
    cm_matrix <- as.matrix(cm$table)

    # Output file path
    filename <- paste0("output/confusion_matrix_", model, "_", title, ".csv")

    # Write confusion matrix to file
    write.csv(cm_matrix, filename, row.names = TRUE)

    # Prepare summary rows (language-level misclassification)
    languages <- unique(model_data$Language)
    model_summary <- list()

    for (lang in languages) {
      lang_data <- model_data[model_data$Language == lang, ]
      total_pro <- sum(lang_data$label == "pro-immigration")
      total_anti <- sum(lang_data$label == "anti-immigration")
      total_tweets <- nrow(lang_data)

      pro_as_anti <- sum(lang_data$label == "pro-immigration" & lang_data$label_llama == "anti-immigration")
      anti_as_pro <- sum(lang_data$label == "anti-immigration" & lang_data$label_llama == "pro-immigration")

      model_summary[[lang]] <- tibble(
        Model = model,
        Dataset = title,
        Language = lang,
        Tweets = total_tweets,
        Pro_as_Anti_Count = pro_as_anti,
        Pro_as_Anti_Perc = ifelse(total_pro > 0, round(100 * pro_as_anti / total_pro, 2), NA),
        Anti_as_Pro_Count = anti_as_pro,
        Anti_as_Pro_Perc = ifelse(total_anti > 0, round(100 * anti_as_pro / total_anti, 2), NA)
      )
    }

    # Combine language-level summaries
    summary_df <- bind_rows(model_summary)

    # Append empty lines, then summary to the same CSV
    write("\n\nLanguage-Level Misclassification Summary:\n", file = filename, append = TRUE)
    suppressMessages(write.table(summary_df, file = filename, sep = ",", row.names = FALSE, append = TRUE, col.names = TRUE))

    # Add to final return object
    results[[model]] <- summary_df
  }

  bind_rows(results)
}
```


```{r}
# Generate accuracy stats for both datasets
accuracy_test <- generate_accuracy_df(df_test, "Test")
accuracy_train <- generate_accuracy_df(df_train, "Train")

# combine into one dataframe
#accuracy_data <- bind_rows(accuracy_test, accuracy_train)
```

```{r}
# save confusion matrix dataset
generate_and_save_cm(df_test, "Test")
generate_and_save_cm(df_train, "Train")
```

```{r}
# dynamic title for the chart
chart_title <- paste("Accuracy Confidence Intervals for", dataset, "Dataset")

# function to create accuracy confidence interval plots
plot_accuracy_intervals <- function(data, title) {
  ggplot(data, aes(x = Language, ymin = Accuracy_Lower, ymax = Accuracy_Upper, color = Model, fill = Model)) +
    geom_errorbar(width = 0.3, position = position_dodge(width = 0.6)) +
    geom_point(aes(y = (Accuracy_Lower + Accuracy_Upper) / 2), size = 3, position = position_dodge(width = 0.6)) +
    scale_x_discrete(labels = function(x) ifelse(x %in% names(language_map), language_map[[x]], x)) +
    labs(title = title,
         x = "Language",
         y = "Accuracy Confidence Interval",
         color = "Model",
         fill = "Model") +
    theme_minimal() +
    theme(legend.position = "bottom",
          panel.background = element_rect(fill = "white", color = NA),
          plot.background = element_rect(fill = "white", color = NA),
          legend.title = element_text(size = 12, face = "bold"),
          legend.text = element_text(size = 10),
          axis.text.x = element_text(angle = 45, hjust = 1))
  
  # save the plot
  ggsave(filename = paste0(chart_title, ".png"),
         path = 'output',
         width = 10, height = 6, dpi = 300)
}

```

```{r}
# plot 'Test' set confidence inervals
chart_title <- "Accuracy Confidence Intervals by Language and Dataset (Test)"
plot_accuracy_intervals(accuracy_test, chart_title)
```

```{r}
# plot 'Train' set confidence inervals
chart_title <- "Accuracy Confidence Intervals by Language and Dataset (Train)"
plot_accuracy_intervals(accuracy_train, chart_title)
```

#### Multilevel Models
```{r}
library(lme4)

# Multilevel logistic regression
logit_mixed <- glmer(Correct ~ Model * Language + label + Train_Test + TranslationQuality + 
             (1 | Language), 
      data = df, family = binomial())

# Summarize the model
summary(logit_mixed)
```
```{r}
# Multilevel logistic regression with random slopes
multilevel_interaction <- glmer(Correct ~ Model + label + Train_Test + TranslationQuality + 
             (1 + Model | Language), 
      data = df, family = binomial())

summary(multilevel_interaction)

```

##### Random Forest
```{r}
tree <- rpart(Correct ~ Model + Language + Train_Test + TranslationQuality + label, # add TrainSize
                         data = df)
rpart.plot(tree, yesno = 2)
```
