---
title: "modelling_llm_performances"
author: "Andrea Nasuto"
date: "2025-04-08"
output: html_document
---

Build a model to assess the performance of a series of LLMs across multiple languages, translations, test/train dataset

The dataset to perform this should have this structure. All variables should be factor/characters:

ID | Model | Language | Translation | Test/Train | Correct
100100 | FT-Multi | IT | 0 | Train | 1
100100 | FT-EN | IT | 1 | Test | 0
100100 | FT-EN-ES | IT | 1 | Train | 1
120300 | LLM2 | EN | EN | Train | 1
...


```{r}
# Set the option to avoid scientific notation
options(scipen = 999)
```

```{r}
# load libraries
library(dplyr)
library(ggplot2)
library(tidyr)
library(purrr)
library(readr)
library(stringr)
library(caret)
library(readxl)
library(pROC)
library(rpart)
library(rpart.plot)
```

```{r}
# dataset paths
base_path_test <- "data/test_dataset_raw"

base_path_train <- "data/train_dataset_raw"

path_translation_test <- "data/test_dataset_translation"

path_translation_train <- "data/train_dataset_translation"

#path_translation_checked_test <- "/Users/Andrea/Library/CloudStorage/OneDrive-TheUniversityofLiverpool/PhD/code/global-immigration-sentiment/validation/tweet_translation_checked/test/"

#path_translation_checked_train <- "/Users/Andrea/Library/CloudStorage/OneDrive-TheUniversityofLiverpool/PhD/code/global-immigration-sentiment/validation/tweet_translation_checked/train/"

# output path
output_dir <- "output"
```

```{r}
# Define dataset directories
dirs <- list(
  list(path = base_path_train, type = "Train"),
  list(path = base_path_test, type = "Test"),
  list(path = path_translation_train, type = "Train"),
  list(path = path_translation_test, type = "Test")
)

# Get all file paths for processing
get_all_files <- function(dirs) {
  all_files <- list()
  for (dir in dirs) {
    if (!is.character(dir$path) || length(dir$path) != 1) {
      stop("Each dir in 'dirs' must contain a character path. Found: ", dir$path)
    }

    # List .csv and .xlsx files
    files <- list.files(dir$path, full.names = TRUE, pattern = "\\.(csv|xlsx)$")

    # Exclude files with 'translation' in the name for base_path_train and base_path_test
    if (dir$path %in% c(base_path_train, base_path_test)) {
      files <- files[!grepl("translation", basename(files), ignore.case = TRUE)]
    }
    
    # Exclude files with 'full' in the name
    files <- files[!grepl("full", basename(files), ignore.case = TRUE)]

    for (f in files) {
      all_files[[length(all_files) + 1]] <- list(file = f, type = dir$type)
    }
  }
  return(all_files)
}
```

```{r}
lang_code_to_name <- function(code) {
  lang_map <- c(
  "mig-ar" = "Arabic",
  "mig-ar-L3" = "Arabic",
  "mig-es" = "Spanish",
  "mig-es-L3" = "Spanish",
  "mig-en" = "English",
  "mig-en-L3" = "English",
  "mig-pt" = "Portuguese",
  "mig-pt-L3" = "Portuguese",
  "mig-fr" = "French",
  "mig-fr-L3" = "French",
  "mig-it" = "Italian",
  "mig-it-L3" = "Italian",
  "mig-de" = "German",
  "mig-de-L3" = "German",
  "mig-hi" = "Hindi",
  "mig-hi-L3" = "Hindi",
  "mig-ko-L3" = "Korean",
  "mig-ko" = "Korean",
  "mig-en-es" = "Eng-Spa",
  "mig-en-es-L3" = "Eng-Spa",
  "mig-multi" = "Multi",
  "mig-multi-L3" = "Multi",
  "mig-mu" = "Multi",
  "mig-mu-L3" = "Multi",
  "mig-hu" = "Hungarian",
  "mig-hu-L3" = "Hungarian",
  "mig-in" = "Indonesian",
  "mig-in-L3" = "Indonesian",
  "mig-es-translation" = "Spanish Translation",
  "mig-es-translation-L3" = "Spanish Translation",
  "mig-tr" = "Turkish",
  "mig-tr-L3" = "Turkish",
  "mig-pl" = "Polish",
  "mig-pl-L3" = "Polish"
  )
  
  if (code %in% names(lang_map)) {
    return(lang_map[[code]])
  } else {
    warning(paste("Unknown or unwanted language code:", code))
    return(NA)
  }
}
```

```{r}
extract_model_dataset <- function(filename) {
  model <- NA
  language <- NA
  
  # Extract model name (e.g., llama-32-3B)
  model_match <- stringr::str_extract(filename, "llama-[^_]+")
  if (!is.na(model_match)) {
    model <- model_match
  } else {
    # Extract model suffix right before _gguf
    model_suffix_match <- stringr::str_match(filename, "_([a-z\\-]+)_gguf")[,2]
    
    if (!is.na(model_suffix_match)) {
      # Map suffix to model name
      if (model_suffix_match == "en-es") {
        model <- "llama-32-3B-en-es-Q4"
      } else if (model_suffix_match == "multi") {
        model <- "llama-32-3B-multi-Q4"
      } else if (model_suffix_match == "en") {
        model <- "llama-32-3B-en-Q4"
      } else if (model_suffix_match == "es") {
        model <- "llama-32-3B-es-Q4"
      } else {
        model <- model_suffix_match  # fallback: just the suffix itself
      }
    }
  }
  
  # Extract full lang key like "mig-fr"
  mig_lang_match <- stringr::str_match(filename, "(mig-[a-z]{2}(?:-L\\d+)?)")[,2]
  
  if (!is.na(mig_lang_match)) {
    language <- lang_code_to_name(mig_lang_match)
  }
  
  # Extract language from _xx_ pattern if language not set
  lang_code_match <- stringr::str_match(filename, "_([a-z]{2})_")[,2]
  if (is.na(language) && !is.na(lang_code_match)) {
    language <- lang_code_to_name(lang_code_match)
  }
  
  return(list(Model = model, Language = language))
}
```

```{r}
process_file <- function(file_path, split_type) {
  # Skip files with "full" or "es-translation"
  if (grepl("full", basename(file_path)) ) {
    return(NULL)
  }
  
  if (!grepl("llama-32-3B", basename(file_path))) {
    return(NULL)
  }
  
  # Read the file based on its extension (CSV or Excel)
  file_extension <- tools::file_ext(file_path)
  if (file_extension == "csv") {
    df <- read_csv(file_path, col_types = cols(.default = "c"))
  } else if (file_extension %in% c("xlsx", "xls")) {
    df <- read_excel(file_path)
  } else {
    return(NULL)  # Skip unsupported file types
  }
  
  # If the data frame is empty or null, return NULL
  if (is.null(df) || nrow(df) == 0) return(NULL)
  
  # Rename "text" to "plaintext" only if "plaintext" doesn't already exist
  if ("text" %in% names(df) && !"plaintext" %in% names(df)) {
    df <- df %>% rename(plaintext = text)
  }

  # Extract model and language information from filename
  meta <- extract_model_dataset(basename(file_path))

  # Set Translation value based on the `translation_check` column in the dataset
  if ("translation_check" %in% colnames(df)) {
    df$TranslationQuality <- ifelse(df$translation_check == 1, 'Good', 'Bad')
    df$Translation <- 1
  } else if (startsWith(file_path, path_translation_train) || startsWith(file_path, path_translation_test)) {
    # If file is from path_translation_train or path_translation_test, set Translation to 1
    df$Translation <- 1
    df$TranslationQuality <- 'Unknown'
  } else {
    # Otherwise, set Translation to 0
    df$Translation <- 0
    df$TranslationQuality <- 'Not Translated'
  }
  
  # Process the dataset
  df_processed <- df %>%
    mutate(
      Model = meta$Model,
      Language = meta$Language,
      Train_Test = split_type,
      Correct = ifelse(label_llama == label, 1, 0),
      file_path = file_path
    ) %>%
    select(message_id, Model, Language, Translation, TranslationQuality, Train_Test, Correct, label, label_llama, plaintext, file_path)

  return(df_processed)
}
```

```{r}
# assign new IDs to each unique plaintext
assign_new_ids <- function(df) {
  df %>%
    group_by(plaintext) %>%
    mutate(newId = cur_group_id()) %>%
    ungroup()
}
```

### Create dataset

```{r}
# Combine the processed data from all files
all_files <- get_all_files(dirs)
list_of_processed_files <- lapply(all_files, function(file_info) {
  process_file(file_info$file, file_info$type)
})
```

```{r}
# create dataset
df <- do.call(rbind, Filter(Negate(is.null), list_of_processed_files))
```

#### Post-processing dataset
```{r}
# assign new ID
df <- assign_new_ids(df)
```

```{r}
# include only quantized at 4bit models
df <- df %>%
  filter(grepl("llama-32-3B", Model) & grepl("Q4", Model))
```

```{r}
# count number of obs per language
df %>%
  count(Language, name = "n_observations") %>%
  arrange(desc(n_observations))

table(df$Language)
```

```{r}
# remove duplicate - each observation should have a unique pair of newId and filepath
df <- df %>%
  distinct(file_path, newId, .keep_all = TRUE) 
```

```{r}
# remove obs where Language is NA
df <- df %>%
  filter(!is.na(Language)) %>%
  distinct(file_path, newId, .keep_all = TRUE)
```

```{r}
# rename models after filtering
model_map <- function(model) {
  if (grepl("llama-32-3B-en-Q4-translation", model)) {
    return("English-Translation")
  } else if (grepl("llama-32-3B-en-es", model)) {
    return("English-Spanish")
  } else if (grepl("llama-32-3B-multi", model)) {
    return("Multilanguage")
  } else if (grepl("llama-32-3B-es", model)) {
    return("Spanish")
  } else if (grepl("llama-32-3B-en", model)) {
    return("English")
  } else {
    return(model)
  }
}
```


```{r}
# Rename models after filtering in df
df$Model <- sapply(df$Model, model_map)
```

```{r}
# remove values where Language is Multi
df <- df %>%
  filter(!(Language %in% c("Multi", "Eng-Spa")))
```

```{r}
# remove all observation where label is NA
df <- df %>%
  filter(!is.na(label))
```

```{r}
# Ensure all variables are factors
df <- df %>%
  mutate(
    message_id = as.character(message_id),
    Model = as.factor(Model),
    Language = as.factor(Language),
    Train_Test = as.factor(Train_Test),
    Translation = as.factor(Translation),
    TranslationQuality = as.factor(TranslationQuality),
    Correct = as.factor(Correct)
  )
```

### Checks
```{r}
# check if there are NAs in df
na_count <- sapply(df, function(x) sum(is.na(x)))
print(na_count)

# check how many correct values are in Correct
table(df$Correct)
```
```{r}
# remove obs with NA in label_llama
df <- df %>%
  filter(!is.na(label_llama))
```

```{r}
df <- df %>%
  group_by(Language, Train_Test) %>%
  mutate(Share_Unrelated = mean(label == "unrelated")) %>%
  ungroup()
```


```{r}
# count number of obs per language
df %>%
  count(Language, name = "n_observations") %>%
  arrange(desc(n_observations))

table(df$Language)
```
```{r}
# count number of unique observations per language
df_unique_counts <- df %>%
  distinct(newId, Language) %>%  # ensure uniqueness of newID per language
  count(Language, name = "n_unique_observations") %>%
  arrange(desc(n_unique_observations))

df_unique_counts
```

```{r}
# count the number of obs with Translation == 'Good'
translation_count <- df %>%
  filter(TranslationQuality == 'Good') %>%
  group_by(Model, Language) %>%
  summarise(count = n())
translation_count
```

### Add language pre-trainign share from Touvron et al. (2023)
```{r}
# Create the language share lookup table
lang_share_lookup <- data.frame(
  code = c("en", "unknown", "de", "fr", "sv", "zh", "es", "ru", "nl", "it", "ja", "pl", "pt", "vi",
           "uk", "ko", "ca", "sr", "id", "cs", "fi", "hu", "no", "ro", "bg", "da", "sl", "hr"),
  Language = c("English", "Unknown", "German", "French", "Swedish", "Chinese", "Spanish", "Russian", "Dutch", "Italian", 
               "Japanese", "Polish", "Portuguese", "Vietnamese", "Ukrainian", "Korean", "Catalan", "Serbian", "Indonesian",
               "Czech", "Finnish", "Hungarian", "Norwegian", "Romanian", "Bulgarian", "Danish", "Slovene", "Croatian"),
  LangShare = c(89.70, 8.38, 0.17, 0.16, 0.15, 0.13, 0.13, 0.13, 0.12, 0.11, 0.10, 0.09, 0.09, 0.08,
                0.07, 0.06, 0.04, 0.04, 0.03, 0.03, 0.03, 0.03, 0.03, 0.03, 0.02, 0.02, 0.01, 0.01)
)

# Merge the LangShare into your existing df by the language code
df <- merge(df, lang_share_lookup, by = "Language", all.x = TRUE)
df$LangShare[is.na(df$LangShare)] <- 0
```

## Model
```{r}
# Set reference levels for the factors
df$Model <- relevel(df$Model, ref = "English") #llama-32-3B-en-Q4 - all the models are 4bit
df$Language <- relevel(df$Language, ref = "English")
df$Train_Test <- relevel(df$Train_Test, ref = "Train")
df$label <- factor(df$label)
df$label <- relevel(df$label, ref = "neutral")
df$TranslationQuality <- factor(df$TranslationQuality)
df$TranslationQuality <- relevel(df$TranslationQuality, ref = "Not Translated") # Not Translated
df$Translation <- factor(df$Translation)

# Fit the logistic regression model
logit_model <- glm(Correct ~ Model + Language + Train_Test + TranslationQuality + label, # add Translated not Translated and Translated(0,1)*TranslationQuality(unknown, bad, good)  
                   data = df, 
                   family = binomial())

# Summarize the model
summary(logit_model)
```

```{r}
# Model 2 - Interactions Model*Language
# Fit the logistic regression model
logit_modelxlanguage <- glm(Correct ~ Model*Language + label + Train_Test + TranslationQuality, # add Translated not Translated and Translated(0,1)*TranslationQuality(unknown, bad, good)  
                   data = df, 
                   family = binomial())

# Summarize the model
summary(logit_modelxlanguage)
```

```{r}
# Model 2 - Interactions Model*Language + Model*Share_Unrelated
# Fit the logistic regression model
logit_modelxShareUnrelated <- glm(Correct ~ Model*Language + Model*Share_Unrelated + Train_Test + TranslationQuality, # add Translated not Translated and Translated(0,1)*TranslationQuality(unknown, bad, good)
                   data = df, 
                   family = binomial())

# Summarize the model
summary(logit_modelxShareUnrelated)
```
###### Visualization of the interaction Model x Language + Model x Share of Unrelated
```{r}
# Define your languages and models
languages <- c("Turkish", "Spanish", "Portuguese", "Polish", "Korean", "Italian",
               "Indonesian", "Hungarian", "Hindi", "German", "French", "Arabic")
models <- c("English-Spanish", "Multilanguage", "Spanish")
shares <- seq(0, 1, 0.1)

# Extract coefficients from the model summary
coefs <- coef(logit_modelxShareUnrelated)
coef_names <- names(coefs)

# Function to get coefficient or return 0 if not found
get_coef <- function(term) {
  if (term %in% coef_names) {
    return(coefs[term])
  } else {
    return(0)
  }
}

# Build the dataset
df_plot <- expand.grid(
  Share_Unrelated = shares,
  Language = languages,
  Model = models
) %>%
  rowwise() %>%
  mutate(
    main_effect = get_coef(paste0("Model", Model)),
    lang_interaction = get_coef(paste0("Model", Model, ":Language", Language)),
    share_interaction = get_coef(paste0("Model", Model, ":Share_Unrelated")),
    total_effect = main_effect + lang_interaction + Share_Unrelated * share_interaction,
    Accuracy = 1 / (1 + exp(-total_effect))  # logistic transformation
  )

# Plot
ggplot(df_plot, aes(x = Share_Unrelated, y = Accuracy, color = Model)) +
  geom_line(size = 1.1) +
  facet_wrap(~ Language, ncol = 4) +
  labs(
    x = "Share of Unrelated Tweets",
    y = "Accuracy", # or Log-Odds
    title = "Model Performance Across Languages and Noise Levels"
  ) +
  scale_x_continuous(labels = scales::percent_format(accuracy = 1)) +
  theme_minimal(base_size = 14) +
  theme(legend.position = "top")

# Save the plot
ggsave(
  filename = "predicted_accuracy_by_language.png",
  plot = last_plot(),
  path = "output",
  width = 12,
  height = 8,
  dpi = 300
)
```


```{r}
# Model 3 - Interactions Model*label
# Fit the logistic regression model
logit_modelxlabel <- glm(Correct ~ Model*label + Language + Train_Test + TranslationQuality, # add Translated not Translated and Translated(0,1)*TranslationQuality(unknown, bad, good)  
                   data = df, 
                   family = binomial())

# Summarize the model
summary(logit_modelxlabel)
```

Idea - to test if the Multilanguage adds noise, we keep the language constant (English) and see how it performs. This means that we need to use only tweets that were classified based on the translation.

The results below seems to show that even when we use only tweets in English, the Multilanguage model performs worse than English-Spanish Model and Spanish Model in classifying unrelated tweets.
This seems to suggest that Multilanguage model is worse when it comes to classifying unrelated tweets, even when we use only tweets in English.

```{r}
# select only tweets classified based on the translation
df_translation <- df %>%
  filter(Translation == 1)

# Fit the logistic regression model
logit_modelxlabel <- glm(Correct ~ Model*label + Language + Train_Test + TranslationQuality, # add Translated not Translated and Translated(0,1)*TranslationQuality(unknown, bad, good)  
                   data = df_translation, 
                   family = binomial())

# Summarize the model
summary(logit_modelxlabel)
```


```{r}
# create log of LangShare
df$LogLangShare <- log1p(df$LangShare)

# Model 4 - Interactions Model*LogLangShare
logit_model_interact <- glm(Correct ~ Model*LangShare + Language + label + Train_Test + TranslationQuality + label,
                            data = df,
                            family = binomial())

summary(logit_model_interact)
```

```{r}
# convert Correct to numeric
df$Correct_Num <- as.numeric(as.character(df$Correct))

lang_summary <- df %>%
  group_by(Language, LangShare) %>%
  summarise(accuracy = mean(Correct_Num, na.rm = TRUE), n = n()) %>%
  ungroup()
```

```{r}
# Show relationship between accuracy and share of language in pretraining phase in LLAMA
# create summary table removing unrelated and English tweets
df_summary <- df %>%
  # remove obs where Language is English
  filter(Language != "English") %>%
  filter(label != "unrelated") %>%
  group_by(Language, Model, LangShare) %>%
  summarise(accuracy = mean(Correct_Num), .groups = "drop")

# plot accuracy by Language and Model
ggplot(df_summary, aes(x = LangShare, y = accuracy, color = Model)) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(title = "Accuracy by Language Share and Model Type",
       x = "Language Share in Pretraining (%)",
       y = "Mean Accuracy") +
  theme_minimal()

# save the plot in output folder
ggsave(file.path(output_dir, "accuracy_by_language_share.png"), width = 10, height = 6)
```

```{r}
# Step 1: Summarize and order Language by mean LangShare
language_order <- df_summary %>%
  group_by(Language) %>%
  summarise(mean_share = mean(LangShare), .groups = "drop") %>%
  arrange(mean_share) %>%
  pull(Language)
# Step 2: Set Language as ordered factor
df_summary$Language <- factor(df_summary$Language, levels = language_order)
```


```{r}
# Plot
ggplot(df_summary, aes(x = LangShare, y = accuracy, color = Language)) +
  geom_point(size = 2) +
  geom_smooth(method = "lm", se = FALSE, color = "black", aes(group = 1)) +
  facet_wrap(~ Model) +
  labs(
    title = "Accuracy vs Language Share by Model",
    x = "Share of Language in LLAMA Pretraining",
    y = "Mean Accuracy",
    color = "Language"
  ) +
  theme_minimal()

# save the plot in output folder
ggsave(file.path(output_dir, "accuracy_vs_language_share_by_model.png"), width = 10, height = 6)
```

```{r}
# Get linear model coefficients per model
coefficients_table <- df_summary %>%
  group_by(Model) %>%
  do(tidy(lm(accuracy ~ LangShare, data = .))) %>%
  filter(term == "LangShare") %>%
  select(Model, Estimate = estimate, Std_Error = std.error, p_value = p.value)

print(coefficients_table)
```


#### Multilevel Models
```{r}
library(lme4)

# Multilevel logistic regression
logit_mixed <- glmer(Correct ~ Model * Language + label + Train_Test + TranslationQuality + 
             (1 | Language), 
      data = df, family = binomial())

# Summarize the model
summary(logit_mixed)
```
```{r}
# Multilevel logistic regression with random slopes
multilevel_interaction <- glmer(Correct ~ Model + label + Train_Test + TranslationQuality + 
             (1 + Model | Language), 
      data = df, family = binomial())

summary(multilevel_interaction)

```

##### Random Forest
```{r}
tree <- rpart(Correct ~ Model + Language + Train_Test + TranslationQuality + label, # add TrainSize
                         data = df)
rpart.plot(tree, yesno = 2)
```
